---
title: "Exercise_5"
author: "Patrick Bigler"
date: "2023-04-24"
output:
  pdf_document: default
  html_document: default
editor_options: 
  markdown: 
    wrap: 75
---

# Report Exercise: "Machine Learning 2"

Course: Applied Geo-data Science at University of Bern (Institute of
Geography)

Supervisor: Prof. Dr. Benjamin Stocker

Adviser: Dr. Koen Hufkens, Pepa Aran, Pascal Schneider

[You have questions to the workflow? Contact the Author:]{.underline}

Author: Bigler Patrick
([patrick.bigler1\@students.unibe.ch](mailto:patrick.bigler1@students.unibe.ch){.email})

Matriculation number: 20-100-178

Reference: Report Exercise 5 (Chapter 10)

```{r Load_packages}
source("../R/general/packages.R")
```

```{r Load_and_read_data}
# Access to the data
url <- "https://raw.githubusercontent.com/geco-bern/agds/main/data/FLX_CH-Dav_FLUXNET2015_FULLSET_DD_1997-2014_1-3.csv"

# Read in the data directly from URL
daily_fluxes.2 <- read.table(url, header = TRUE, sep = ",")

# Write a CSV file in the respiratory
write_csv(daily_fluxes.2, "../data/re_ml_02/daily_fluxes.csv")
```

```{r Read_the_data}
# Read the file
daily_fluxes <- read_csv("../data/re_stepwise/FLX_CH-Lae_FLUXNET2015_FULLSET_HH_2004-2006_CLEAN.csv")
```
