---
title: "Report Exercise: re_ml_02"
author: "Patrick Bigler"
date: "2023-04-24"
output:
  html_document:
    toc: yes
    toc_float: yes
    number_sections: yes
editor_options: 
  markdown: 
    wrap: 75
---

Course: Applied Geo-Data Science at the University of Bern (Institute of
Geography)

Supervisor: Prof. Dr. Benjamin Stocker

Adviser: Dr. Koen Hufkens, Pepa Aran, Pascal Schneider

Further information: <https://geco-bern.github.io/agds/>

[Do you have questions about the workflow? Contact the Author:]{.underline}

Author: Bigler Patrick (patrick.bigler1\@students.unibe.ch)

Matriculation number: 20-100-178

Reference: Report Exercise 7 (Chapter 10)

# Introduction

## Objectives

In this exercise we explore the role of structure in the data for model
generalisability and how to best estimate a "true" out-of-sample error that
corresponds to the prediction task. The task here is to train a model on
ecosystem flux observations from one site and predict them for another site
(spatially upscaling). With a realistic problem, the following goals should
be trained:

-   Concept of cross validation

-   spatially up-scalling

## Theory^1^

The theory is very similar to re_ml_01. We change only our method to
cross-validation. Validation is not the same as testing the data.
Validation is a concept that will be applied during the data training.

Cross-validation (cv) is a resampling method that randomly divides the
training data into n-folds of approximately equal size. We fit the model
for n-1 folds and use the remaining fold to compute model performance. We
repeat that n times, and for each time we use a different fold as a
validation set. After we finished this procedure, we had n estimates of the
generalization error. Thus, the n-fold cv estimate is computed by averaging
the n test errors, providing us with an approximation of the error we might
expect on unseen data. There is no formal rule as to the size of *n*.
However, as *n* gets larger, the difference between the estimated
performance and the true performance to be seen on the test set will
decrease. But we will need more time to compute the models.

For KNN, *k* is the only hyper-parameter. In exercise re_ml_01, we
demonstrated how to find *optimal k*. If we train our data with the
cv-method, we get a model that has a better bias-variance trade-off than
the models before. With *optimal k* and cross-validation, we have a "true"
out-of-sample prediction. That is why we implemented a cv-KNN algorithm.

# Method

We split our data and use 80% for training and 20% for model evaluation. We
use 10 folds for cross-validation. Because the algorithm uses distance, we
have to standardize our data. With the box-cox function, we transform our
data into a normal distribution. We use set.seed for reproducibility.

## R: Evaluation of the data

The open-source program R-Studio (Version 2022.12.0+353) was used for all
studies. The data handling in this program is done using the R language. We
utilize the R-package "Tidyverse" and other packages to process the data
effectively.

# Programming and data evaluation

## Packages

The following code chunk contains all packages we need. Important is the
package "conflicted". It enables us to choose if different functions have
the same call, but do not make the same thing (a function in a certain
package can have the same name as a function from another package). In this
case, we will set preferences.

```{r Load_packages, error=FALSE, message=FALSE, warning=FALSE}
source("../../R/general/packages.R")
```

## Read the file

First, we get the data and save it in our repository. But we must do this
only once. If the file exists, then we can read it directly. That is why we
implemented an if-else statement. We know, that we can do easier. But we
want to demonstrate, that there are many ways to read the file.

```{r Load_and_read_data, error=FALSE, message=FALSE, warning=FALSE}
name.of.file <- "../../data/re_ml_02/daily_fluxes.davos.csv"

# If do not exists such a file, create it!
if (!file.exists(name.of.file)){
  
  # Access to the data
  url.1 <- "https://raw.githubusercontent.com/geco-bern/agds/main/data/FLX_C
  -Dav_FLUXNET2015_FULLSET_DD_1997-2014_1-3.csv"
  
  # Read in the data directly from URL
  daily_fluxes.davos <- read.table(url.1, header = TRUE, sep = ",")
  
  # Write a CSV file in the respiratory
  write_csv(daily_fluxes.davos, "../../data/re_ml_02/daily_fluxes.davos.csv")
  
  # Read the file
  daily_fluxes.davos <- read_csv("../../data/re_ml_02/daily_fluxes.davos.csv")
  
  # If exists such a file, read it only!
  }else{daily_fluxes.davos <- read_csv("../../data/re_ml_02/daily_fluxes.davos.csv")}

name.of.file <- "../../data/re_ml_02/daily_fluxes.laegern.csv"

# If do not exists such a file, create it!
if (!file.exists(name.of.file)){
  
  # Access to the data
  url.2 <- "https://raw.githubusercontent.com/geco-bern/agds/main/data/FLX_CH
  -Lae_FLUXNET2015_FULLSET_DD_2004-2014_1-4.csv"
  
  # Read in the data directly from URL
  daily_fluxes.davos <- read.table(url.1, header = TRUE, sep = ",")
  
  # Write a CSV file in the respiratory
  write_csv(daily_fluxes.laegern, "../../data/re_ml_02/daily_fluxes.laegern.csv")
  
  # Read the file
  daily_fluxes.laegern <- read_csv("../../data/re_ml_02/daily_fluxes.laegern.csv")
  
  # If exists such a file, read it!
}else{daily_fluxes.laegern <- read_csv("../../data/re_ml_02/daily_fluxes.laegern.csv")}
```

## Data cleaning

We can see that some columns contain -9999 as a value. Our quality function
changed that to NA. Then we use ymd() from the "lubridate" package to
rewrite the date in a proper way. Further, we want only columns that
contain good quality. For that, we check selected columns against their
quality control columns. If the proportion of good measurement is less than
80%, then we overwrite the value with NA (and do not drop the row!). Now we
know that our data has high quality, and we can perform our analyses with
it.

Here, we work with two locations (Davos and Lägern). But we do not want to
do everything twice. That is why we define a new data frame that contains
all the information about the two locations. We use .id = "id" because we
must know where the values come from. But this is not so elegant. We change
the column name (id = Location) and replace the values (1 = Davos, 2 =
Lägern). If we do that, we can use the filter function from the package
"dplyr" to efficiently analyze our data.

We can see that there are many missing values in incoming long-wave
(LW_IN_F) in Davos and precipitation (P_F) in Lägern. We would lose almost
half of our data in our recipe call (data training). That is why we do not
use this column. After we descard the columns we make a quality check again
and we see: our data is ok! We bind our rows and make a quality check one
last time. Now we are sure, that everything is ok.

```{r Dataquality, error=FALSE, message=FALSE, warning=FALSE}
# Load the function in the file (we use the function from another markdown again)
source("../../R/re_ml_01/function.use.good.quality.only.R")
source("../../R/general/function.visualize.na.values.R")

# Function call to clean the data. But the quality is poor and we do not use them!
daily_fluxes.dav.no <- use.good.quality.only(daily_fluxes.davos)
daily_fluxes.lag.no <- use.good.quality.only(daily_fluxes.laegern)

# Visualize the result ---> many NA! We would loos to much data.
visualize.na.values.without.groups(daily_fluxes.dav.no, c(" Davos: Problem with LW_IN_F!"))
visualize.na.values.without.groups(daily_fluxes.lag.no, c(" Lägern: Problem with P_F!"))

# Function call to clean the data (We use another function to clean our data)
daily_fluxes.dav <- use.good.quality(daily_fluxes.davos)
daily_fluxes.lag <- use.good.quality(daily_fluxes.laegern)

# Visualize the results --> Now we can work!
visualize.na.values.without.groups(daily_fluxes.dav, c("Davos: Everything ok!"))
visualize.na.values.without.groups(daily_fluxes.lag, c("Lägern: Everything ok!"))

# We create a new data frame with bind_rows. We us "id" as an identifier. 
daily_fluxes_both <- bind_rows(daily_fluxes.dav, daily_fluxes.lag, .id = "id")

# We change the ID in our location names
daily_fluxes_both <- daily_fluxes_both|>
  rename("Location" = "id")|>
  mutate(Location = ifelse(Location == 1, "Davos", "Lägern"))

# Function call: we make a quality control one last time --> everything is ok!
visualize.na.values.without.groups(daily_fluxes_both, c("Pooled: Everything ok!"))
```

## Split the data

Here, we split our data into a training set and a test set. First, we will
split the data (80% training and 20% testing). We set the seed for a
pseudo-random choice (for reproducibility).

```{r error=FALSE, message=FALSE, warning=FALSE}
# For reproducibility (pseudo-random)
set.seed(123)  
# Split 80 % to 20 % 
split_davos <- rsample::initial_split(daily_fluxes_both|> 
                                        dplyr::filter(Location == "Davos"), 
                                        prop = 0.8, strata = "VPD_F")

split_laegern <- rsample::initial_split(daily_fluxes_both|> 
                                          dplyr::filter(Location == "Lägern"), 
                                          prop = 0.8, strata = "VPD_F")

split_both <- rsample::initial_split(daily_fluxes_both, prop = 0.8, strata = "VPD_F")

# Split Davos
daily_fluxes_davos_train <- rsample::training(split_davos)
daily_fluxes_davos_test <- rsample::testing(split_davos)

# Split Lägern
daily_fluxes_laegern_train <- rsample::training(split_laegern)
daily_fluxes_laegern_test <- rsample::testing(split_laegern)

# Split pooled
daily_fluxes_both_train <- rsample::training(split_both)
daily_fluxes_both_test <- rsample::testing(split_both)
```

## Find optimal k

We use a sequence to determine the *optimal k*. The model with an *optimal
k* has the smallest mean absolute error (MAE). But we are also interested
in other metrics, like RSQ. Our function will determine RSQ as well. We
create three models:

1.  We train a model with data from Davos
2.  We train a model with data from Lägern
3.  We train a model with data from Davos+Lägern (pooled)

### Davos

We can see that the *optimal k* is 50. This seems odd because the data are
the same as for the exercise re_ml_01 and there was our *optimal k* = 15.
But we have to consider that we used a different approach to create our
model. We use "cross-validation," which is why we find different *optimal
k* for the same data (and the same set.seed). Further, in this exercise we
us 80% instead of 70% of the data set for training.

We also used other sequences to find the *optimal k* (because with the call
below we only know *optimal k* lies in a certain interval). Only the
results are presented here. We set eval = false because the time to compute
the value is high. If you want to check, please run the code manually (it
will work).

```{r optimal_k_davos, eval=FALSE, error=FALSE, message=FALSE, warning=FALSE}
source("../../R/re_ml_02/function.parameter.extracter.cv.R")

# Define a sequence for k. Use 1,2,3,4 to show the curve at the beginning
my.sequence <- c(1, 2, 3, 4, seq(5, to = 100, by = 5))

# Visualize the MAE and RSQ --> optimal k is 50, note that 10 is the number of n-folds
parameter.extracter.knn.cv(my.sequence, 
                           daily_fluxes_davos_train, 
                           daily_fluxes_davos_test, 10)
```

### Lägern

We make the same thing for Lägern. Here is *optimal k* equal to 23. We set
eval = false because the time to compute the value is high. If you want to
check, please run the code manually (it will work).

```{r optimal_k_laegern, eval=FALSE, error=FALSE, message=FALSE, warning=FALSE}
source("../../R/re_ml_02/function.parameter.extracter.cv.R")

# Define a sequence for k. Use 1,2,3,4 to show the curve at the beginning
my.sequence <- c(1, 2, 3, 4, seq(5, to = 100, by = 5), 23)

# Visualize the MAE and RSQ --> optimal k is 23 (between 20:30)
parameter.extracter.knn.cv(my.sequence, 
                           daily_fluxes_laegern_train, 
                           daily_fluxes_laegern_test, 10)
```

### Davos + Lägern

We do the same thing for the pooled data, and we find an *optimal k* equal
to 49. Please be patient. The calculation time is high.

```{r optimal_k_pooled, error=FALSE, message=FALSE, warning=FALSE}
source("../../R/re_ml_02/function.parameter.extracter.cv.R")

# Define a sequence for k. Use 1,2,3,4 to show the curve at the beginning
my.sequence <- c(1, 2, 3, 4, seq(5, to = 100, by = 5),49)

# Visualize the MAE and RSQ --> optimal k is 25 (between 20:30)
parameter.extracter.knn.cv(my.sequence, 
                       daily_fluxes_both_train, 
                       daily_fluxes_both_test, 10)
```

## Best models

Now we know the *optimal k* for each model, and we recalculate them.

```{r error=FALSE, message=FALSE, warning=FALSE}
source("../../R/re_ml_02/function.knn.cv.model.R")

# Function call for Davos, 50 = k, 10 = cv
knn.model.davos.optimal <- knn.cv.model(daily_fluxes_davos_train, 50, 10)
# Function call for Lägern, 23 = k, 10 = cv
knn.model.laegern.optimal <- knn.cv.model(daily_fluxes_laegern_train, 23, 10)
# Function call for Davos and Lägern, 49 = k, 10 = cv
knn.model.both.optimal <- knn.cv.model(daily_fluxes_both_train, 49, 10)
```

### Visualization of the models

We evaluate our models. For that, we compare each train data to all other
test data. It follows, that we need 9 evaluations.

```{r error=FALSE, message=FALSE, warning=FALSE}
# Load the function
source("../../R/re_ml_01/function.evaluation.model.R")

# Model Davos
P_1 <- eval_model(knn.model.davos.optimal, daily_fluxes_davos_train, 
                  daily_fluxes_davos_test, 
                  c("Davos (opt. k = 50)"), c("Davos"))

P_2 <- eval_model(knn.model.davos.optimal, daily_fluxes_davos_train, 
                  daily_fluxes_laegern_test, 
                  c("Davos (opt. k = 50)"), c("Lägern"))

P_3 <- eval_model(knn.model.davos.optimal, daily_fluxes_davos_train, 
                  daily_fluxes_both_test, 
                  c("Davos (opt. k = 50)"), c("Davos and Lägern"))

# Model Lägern
P_4 <- eval_model(knn.model.laegern.optimal, daily_fluxes_laegern_train, 
                  daily_fluxes_davos_test, 
                  c("Lägern (opt. k = 23)"), c("Davos"))

P_5 <- eval_model(knn.model.laegern.optimal, daily_fluxes_laegern_train, 
                  daily_fluxes_laegern_test,
                  c("Lägern (opt. k = 23)"), c("Lägern"))


P_6 <- eval_model(knn.model.laegern.optimal, daily_fluxes_laegern_train, 
                  daily_fluxes_both_test, 
                  c("Lägern (opt. k = 23)"), c("Davos+Lägern"))

# Model Davos und Lägern
P_7 <- eval_model(knn.model.both.optimal, daily_fluxes_both_train,
                  daily_fluxes_davos_test, c("Davos and Lägern (opt. k = 49)"), 
                  c("Davos"))

P_8 <- eval_model(knn.model.both.optimal, daily_fluxes_both_train, 
                  daily_fluxes_laegern_test, c("Davos and Lägern (opt. k = 49)"), 
                  c("Lägern"))


P_9 <- eval_model(knn.model.both.optimal, daily_fluxes_both_train,
                  daily_fluxes_both_test, c("Davos and Lägern (opt. k = 49)"), 
                  c("Davos+Lägern"))

# Davos: Within side, across side and both locations
cowplot::plot_grid(P_1, P_2, P_3, ncol = 1)

# Lägern: Within side, across side and both locations
cowplot::plot_grid(P_4, P_5, P_6, ncol = 1)

# Pooled: Within side, across side and both locations
cowplot::plot_grid(P_7, P_8, P_9, ncol = 1)
```

## Residue

We want more information about the residue. We use a flexible function to
get this information. We plot a time variation. Now we can analyze whether
there are any patterns. We use the pooled KNN-model to predict the
locations because we are interested in a very generalized model. We
visualize the whole data set (all observations and all test data), and we
can see that there is annual cycle for both locations.

```{r error=FALSE, message=FALSE, warning=FALSE}
# Load the function into the file
source("../../R/re_ml_02/function.plot.function.R")

# We plot the actually measured GPP in the pooled data set.
plot.time.variation(daily_fluxes_both,
     daily_fluxes_both$TIMESTAMP,
     daily_fluxes_both$GPP_NT_VUT_REF,c("Overview about the pooled measurements"), 
     sub.lab = c("optimal k = 49"),prediction = FALSE)

# We plot the test data set from the pooled data set. We did not do any prediction
plot.time.variation(daily_fluxes_both_test,
     daily_fluxes_both_test$TIMESTAMP,
     daily_fluxes_both_test$GPP_NT_VUT_REF, c("Overview about the pooled test data"),
     prediction = FALSE, sub.lab = c("optimal k = 49"))
```

```{r error=FALSE, message=FALSE, warning=FALSE}
# Load the function into the file
source("../../R/re_ml_02/function.plot.function.R")

# We plot the predicted test data set. 
plot.time.variation(daily_fluxes_both_test,
     daily_fluxes_both_test$TIMESTAMP,
     daily_fluxes_both_test$GPP_NT_VUT_REF,c("Overview about the predicted pooled test data"),
     prediction = TRUE, mod1 = knn.model.both.optimal, sub.lab = c("optimal k = 49"))
```

### Residue for Davos

First, we take a look at Davos. We can see that the model underestimates
the GPP values. But more important is the fact, that there is a annual
cycle. In summer, the model overestimate the GPP and in winter the model
underestimate GPP. We want to know, where exactly and why. To answer this
questions, we push our model to a higher resolution. We take a closer look
and plot the residue as a function of the day of the year. Now we are sure
that our model has a negative bias for the winter season and a positive
bias for the summer.

```{r error=FALSE, message=FALSE, warning=FALSE}
# Load the function
source("../../R/re_ml_02/function.vis.residue.R")

# Residue for Davos: train: pooled, test: Lägern (year)
time.variation.year(knn.model.both.optimal, daily_fluxes_davos_test,
                    c("Train: Pooled, Test: Davos, k = 49"), 
                    c("re_ml_2 (Chapter 10)"))

# We want a better resolution of the residue of Davos (daily)
time.variation.year(knn.model.both.optimal, daily_fluxes_davos_test,
                    c("Train: Pooled, Test: Davos, k = 49"), 
                    c("re_ml_2 (Chapter 10)"), annual = FALSE)
```

### Residue for Lägern

We do the same thing for Lägern. We can also see that there is an annual
cycle. The model overestimates the GPP values during the summer and
underestimates them during the winter. To explain that, we need a higher
resolution. We plot the residue as a function of the day of the year.

```{r error=FALSE, message=FALSE, warning=FALSE}
# Load the function
source("../../R/re_ml_02/function.vis.residue.R")

# Residue for Davos: train: pooled, test: Lägern (year)
time.variation.year(knn.model.both.optimal, daily_fluxes_laegern_test,
                    c("Train: Pooled, Test: Lägern, k = 49"), 
                    c("re_ml_2 (Chapter 10)"))

# We want a better resolution of the residue of Lägern (daily)
time.variation.year(knn.model.both.optimal, daily_fluxes_laegern_test,
                    c("Train: Pooled, Test: Lägern, k = 49"), 
                    c("re_ml_2 (Chapter 10)"), annual = FALSE)
```

# Discussion

## Location

We have merged two data sets. But the two locations are very different.
Table 1 give you a overview:

**Table 1:** Comparison of the measuring stations (Davos vs. Lägern)

| Quantity             | Davos                                                                                                                                                                                                          | Lägern                                                                                                                                                                                                                                                       |
|:------------------|:--------------------------|:-----------------------------|
| Coordinates^2^:      | 46.81°N, 9.84°E                                                                                                                                                                                                | 47.48°N, 8.4°E                                                                                                                                                                                                                                               |
| Elevation^2^:        | 1639 m.a.s.l.                                                                                                                                                                                                  | 689 m.a.s.l.                                                                                                                                                                                                                                                 |
| Exposure^2^:         | South-East slope                                                                                                                                                                                               | Summit                                                                                                                                                                                                                                                       |
| 2m Temperature^3^    | 2.8° C                                                                                                                                                                                                         | 8.3° C                                                                                                                                                                                                                                                       |
| Precipitation^3^     | 1062 mm per year                                                                                                                                                                                               | 1100 mm per year                                                                                                                                                                                                                                             |
| Climate^3^ [Koeppen] | Tundra                                                                                                                                                                                                         | \-                                                                                                                                                                                                                                                           |
| Hours of sunshine^2^ | 1537 h - 2076 h                                                                                                                                                                                                | 1462 h - 2152 h                                                                                                                                                                                                                                              |
| Wind^2^              | 8 km/h                                                                                                                                                                                                         | 17 km/h                                                                                                                                                                                                                                                      |
| Vegetation^3^:       | ENF (Evergreen Needleleaf Forests: Lands dominated by woody vegetation with a percent cover \>60% and height exceeding 2 meters. Almost all trees remain green all year. Canopy is never without green foliage | MF (Mixed Forests: Lands dominated by trees with a percent cover \>60% and height exceeding 2 meters. Consists of tree communities with interspersed mixtures or mosaics of the other four forest types. None of the forest types exceeds 60% of landscape.) |

## Model

We tried different approaches to modeling GPP. We took some metrics and
hyper-parameter into account (e.g., RSQ, MAE, *k*). Our goal was to create
a model that is generalizable. That means that we want a model that can
predict GPP in different locations (Davos or Lägern). After we found the
probably best model, we applied the model to predict the test data. The
difference between predicted test data and test data is called the bias.
With this parameter, we are able to discuss our analysis.

### How do the model metrics on the test set compare to the true out-of-sample setup above?

We can see that if we train our data with cross-validation, our true
out-of-sample performed best on the test data from the same location. If we
predict another location, the model performs worst. If we predict pooled
data, the model performs better. That is not surprising at all. We take a
closer look:

[Davos:]{.underline}

If we train a model with data from Davos and optimize it (*k* = 50), RSQ is
0.73 (which is actually very good). The RMSE is 1.43. If we predict the
test data from Davos, the RSQ is 0.73 and the RMSE is 1.44. The bias is
about -0.06 which means our model underestimate gpp in general. If we
predict Lägern, RSQ is decreasing to 0.52, and RMSE is increasing to 3.24.
It follows that we cannot predict Lägern with a Davos model. If we predict
pooled data, RMSE is decreasing to 0.61 and increasing to 2.24. The bias is
about 0.42.

[Lägern:]{.underline}

For Lägern, the situation is quite similar to Davos. If we train a model
with Lägern data and optimize it (*k* = 23), RSQ is 0.66 and RMSE is 2.44.
The bias is 0.01 which means the model overestimate GPP in general. We also
see that if we predict Davos, RSQ decreasing to 0.55 and RMSE increasing to
2.33. The bias increase to -1.07. If we predict pooled data, RSQ decreases
to 0.61 and RMSE increases to 2.32. The bias is about -0.68 and therefore
higher than for the Davos model (but opposite sign!).

[Pooled:]{.underline}

Our goal is to create a model that can predict different locations. To
achieve this, we trained a model and optimized it (*k* = 38). If we predict
pooled data, RSQ is 0.67, and RMSE is 2. The bias is almost zero. If we
predict Davos, RSQ increases to 0.71, RSME decreases to 1.53, and the bias
to about -0.29. This is remarkable and surprising. This is almost the same
model performance as the Davos model. If we predict Lägern, RSQ decreases
to 0.64, RMSE increases to 2.62, and the bias to 0.59.

We also see, that for Davos, the model overestimates GPP with
almost $2\:\mu mol*m^{-2}*s^{-1}$ during summer-season and underestimate
GPP with almost $-2\:\mu mol*m^{-2}*s^{-1}$ during winter-season. For all
other seasons, the mean of the prediction is almost zero. For Lägern, the
model underestimates GPP with almost $5\:\mu mol*m^{-2}*s^{-1}$ during
summer and almost $-5\:\mu mol*m^{-2}*s^{-1}$ during winter. It seems that
the model performs way better for Davos than for Lägern. The reason could
be that we have a lot more data for Davos (6574 rows) than for Lägern (4018
rows).

### Interpret differences. Is it a valid approach to perform model training like this? 

Before we interpret the difference, we must know what we want to do. Are we
only interested in a specific location? Or is our interest in predicting
GPP in general?

Let's assume we are interested in a location. It is obvious that we get a
better model if we use only data from the location of interest. But often
we are not interested in a specific location, and if we were, then we must
have something to compare with. Therefore, we need a generalized model to
predict different locations.

Of course, we are not yet satisfied with the model. We suggest that the
project be extended to the whole of Switzerland. Then we would have dozens
of measuring stations on different terrains. We could then classify the
stations. For example, according to their exposure. Then a model could be
created for all classes. For example, the model "summit" was trained with
data from stations at summits. It could also make sense to divide the
stations into regions or altitude levels. One has to make a trade-off
between generalizability and accuracy. If we generalize, we have fewer
models. However, this increases the error. If we do not generalize, then we
have higher accuracy, but we have to use a different one for each location.

### Interpret biases of the out-of-sample predictions with a view to the site characteristics

A very important parameter is the bias. We can see that our model
underestimates GPP for Davos and underestimates for Lägern. The reason for
that is not clear. In Davos, the model overestimate GPP about 150th day of
the year (mid May). In Lägern this happens already for the 100th day of the
year (mid April). It could be the elevation (see table 1). In April, Davos
has still snow and the vegetation growth later. It also could be
temperature because low temperature inhibit GPP.

Another explanation could be exposure (see table 1). In Lägern, the station
is at a "summit". That means there is more wind. Maybe the wind falsifies
the measurement (but we do not think that because that would be very
unprofessional). It is different whether the sun shins direct on the
surface or is there always a shadow (different amount of energy per square
meter).

In summary, we can say that we should not pool the data randomly. It is
important that we compare same with same. We suggest, that we classify the
measurement stations because topology, meteorological parameters or
elevation influence the model.

# Bibliography

[(1) Cross-validation]{.underline}

<https://bradleyboehmke.github.io/HOML/process.html#k-fold-cross-validation>

[(2) Meteorological Data]{.underline}

Data for Davos:
<https://www.meteoschweiz.admin.ch/service-und-publikationen/applikationen/messwerte-und-messnetze.html#param=messnetz-automatisch&lang=de&station=DAV&chart=year>

Data for Lägern:
<https://www.meteoschweiz.admin.ch/service-und-publikationen/applikationen/messwerte-und-messnetze.html#param=messnetz-automatisch&lang=de&station=LAE&chart=year>

[(3) Fluxnet]{.underline}

<https://fluxnet.org/sites/siteinfo/CH-Dav>
