---
title: "Report_Exercise_2"
author: "Patrick Bigler"
date: "2023-04-10"
output: html_document
editor_options: 
  markdown: 
    wrap: 75
---

# Report Exercise: "Data Visualization"

Course: Applied Geo-data Science at University of Bern (Institute of
Geography)

Supervisor: Prof. Dr. Benjamin Stocker

Adviser: Dr. Koen Hufkens, Pepa Aran, Pascal Schneider

You have questions to the workflow? Contact the Author:

Author: Bigler Patrick (patrick.bigler1\@students.unibe.ch)

Matriculation number: 20-100-178

Reference: Report Exercise 2 (Chapter 4)

## Introduction and Hypothesis

Air quality depends on factors such as particulate matter (PM10 or PM2.5),
nitrogen dioxide $(NO_2)$ or ozone $(O_3)$. In this exercise, only the
tropospheric ozone is considered for air quality. Ozone is not a health
hazard per se but a high concentration over a longer period of time can
affect the health (Brönnimann 2013, 23). But what does that mean? For
Switzerland, the threshold value is $120\: \frac{\mu g}{m^3}$. This value
is the average for an hour and can be exceeded only once per year (it
exists other definitions for a threshold, but here we use this definition
(Brönnimann 2013, 23). However, this unit depends on the air pressure,
which decreases with altitude. The volume mixing ratio with the unit parts
per million is more suitable which is about 60 ppb which we use in this
exercise. (Brönnimann 2013, 26).

Ozone is a molecule consisting of 3 oxygen atoms and has various physical
(absorption of UV and IR radiation) and chemical properties (high
reactivity). Ozone emerges due to a 2 phase chemical reaction (Brönnimann
2013, 75):

1.  $NO_2 + h\nu \left(\lambda < 420 nm \right) \rightleftharpoons NO + O$

2.  $NO + O_3 \rightleftharpoons NO_2 + O_2$

From the reaction equations it follows that UV radiation, oxygen and
nitrogen dioxide are needed for the formation of ozone . The first two are
available in almost unlimited quantities. Thus, the production depends
mainly on nitrogen dioxide. For the planetary boundary layer (and therefore
for the surface as well), the motorized traffic is the main source
(Brönnimann 2013, 95). Lastly, many chemical reactions are temperature
dependent. This is also true for ozone. The higher the temperature, the
more the balance of the reaction shifts towards ozone production and the
concentration increases (Brönnimann 2013, 95).

In summary, the ozone concentration depends probably on temperature (shift
of the balance), solar radiation (mostly UV-radiation) and the
concentration of nitrogen dioxide (which depends on traffic and wind). With
these considerations, a data set on air-quality can be studied with a focus
on ozone.

This exercise is supposed to answer the following research question:

*"Quantify the dependency of ozone concentration from temperature, solar
radiation and wind"*

## Methods

The method is quite straigth forward.

1.  We make some quality control calls
2.  We check the source because the data set is a second source

First, there is a little bit of data quality control. We have to know, what
variables contains the data set and how many missing values are there.
After that we check the sources and decide if we want to use the data or if
there any changes necessary. Then we make some statistic calls and extract
and visualize the main statistic parameters. With this knowledge we will be
able to describe the ozone with selected variables.

## Programming and data evaluation

## Packages

The following code chunk contains all packages we need. Important is the
package "conflicted". It enables us to chose if different functions have
the same call but do not make the same thing (a function in a certain
package can have the same name as a function from another package).

```{r activate_the_packages_which_are_required}
source("../R/general/packages.R")
```

### Original source of the file

The data is example data frame in the R package "datasets". First, the data
frame is a secondary source an we must know about the contain, the values,
the units and, of course, the primary sources. For that we can use R to
gather all these information. With help() we see that the varibales do not
have the SI-units. We also check the NAs. With str() we can see that there
are NAs. We call a function to visualize the NAs.

```{r shows_the_data_frame_airquality}
# Read the data
airquality <- datasets::airquality

# Ask for details (units, source and references)
help("airquality")

# check data type of every variable in data frame
str(airquality)

# Load the function
source("../R/general/function.visualize.na.values.R")

# Function call
visualize.na.values(airquality)
```

The data frame shows different variables at different times. Start was
05.01.1973 (a Tuesday) and ends at 09.30.1973 and each day has an entry (it
could be "NA" as well). We distinguish the variable according to the source
and describing them.

Data for ozone were collected from the New York State Department of
Conservation.

-   Ozone [ppb] represent the average of the concentration between 1pm -
    3pm at Roosevelt Island, NYC.

Data for the meteorological data were collected from the US. National
Weather Service.

-   Wind [mph] is calculated as the average wind speed at 7am and 10am at
    La Guardia Airport, NYC

-   Temperature [degrees Fahrenheit] is the maximum daily temperature at la
    Guardia Airport, NYC

-   Solar. R [Langley] average in the frequency band 4000-7700 Angstroms
    (wavelength: 400-770 nano meter) from 8am to noon in Central Park NYC.

### Editing the file in SI-Units

It is strongly recommended to use the SI unit system because it is
simplifies the reproducibility and the interpretation and discussion of the
data. It follows the formulas to calculates the values from there units
into the SI-units system

-   For temperature the following formula was used:
    $$\frac{x \left[°Fahrenheit \right]- 32}{1.8} +273.15 \Longleftrightarrow x\left[Kelvin \right]$$

-   For wind speed the following formula was used :
    $$1 \left[\frac{mile}{hour} \right] \Longleftrightarrow 0.44704 \left[\frac{meter}{second}\right]$$

-   For solar radiation the following formula was used :
    $$1 \left[ Langley \right] \Longleftrightarrow 41840 \left[ \frac{Joule}{m^2}\right]$$

```{r editing_the_data_frame_in_SI_units}
# Load the function
source("../R/airquality/function.unit.converter.R")

# Function call
airquality_mutated <- unit.converter(airquality)
```

## Statistics

Now that all units have been converted to SI units, the question about the
position, scattering and distribution parameters arises. For the ozone
concentration is the median the right position parameter because it is
robust against outliers. This is important because a short time higher
concentration of ozone is probably not a problem for the human well-being
but for a long term it could be. If we use the mean, we loss the overall
picture due to a big one time outlier. If the median has been used it
should be used the IQR as a scattering parameter. But we are also
interested in the maximal values. A environment with a very high ozone
concentration can be toxic (even a short time exposure).

The data shows us that the monthly median (may = 18 ppb, Jun = 23 ppb, Jul
= 60 ppb, Aug = 52 ppb, Sept = 23 ppb) is lower or equal to the threshold
value (60 ppm). Therefore, there the air-quality is acceptable. But there
are measurements in every month which exceed the threshold value (.75
quantile, maximum ). That means that there are isolated days when the ozone
concentration is problematic. If we compare the span (167 ppb) with the IQR
(45.25 ppb) we notice that the IQR is clearly smaller than the span. That
means that we indeed have outliers but the the data core is small.

```{r statistical_parameters}
# Monthly median for temperature, ozone, solar and wind
airquality_mutated |>
# Calculate the median for the ozone concentraion  
  summarise(median.temp = median(Temp, na.rm = TRUE),
            median.wind = median(Wind, na.rm = TRUE),
            median.ozone = median(Ozone, na.rm = TRUE),
            Min.ozone = min(Ozone, na.rm = TRUE),
            Q1.ozone = quantile(Ozone, 0.25, na.rm = TRUE),
            Q2.ozone = quantile(Ozone, 0.5, na.rm = TRUE),
            Q3.ozone = quantile(Ozone, 0.75, na.rm = TRUE),
            Max.ozone = max(Ozone, na.rm = TRUE),
            IQR.ozone = IQR(Ozone, na.rm = TRUE),
            Span = max(Ozone, na.rm = TRUE) - min(Ozone, na.rm = TRUE),
            Average.ozone = mean(Ozone, na.rm = TRUE))
```

We can see the same in boxplots. But in a box-whisker plot are some
information hidden because its shows only the percentiles. Therefore, a
jitters plot is added. The plot shows us that there are outliers, but the
boxes for May, June and September are small which means the data scattering
is small. The situation for July and August is different. The large box
indicates a large IQR for these months and the data scattering is
relatively high. Additional, the plot shows also the mean value. It is
higher than the median for each box. Therefore, the skewness is for all
months positive.

```{r boxplot_ozone_per_month}
# Load the function
source("../R/airquality/function.boxplot.R")

# Function call
airquality.boxplot(airquality_mutated)
```

The density plot shows us that the data have a big span. But the core of
the data (area with a high density) has an acceptable span (which is
basically the IQR). The graphs shows us also that the density for the
months May, June and September is bi-modal. That means, that there are
outliers (which we already see in the boxplot). July and August are
mono-modal but have a large span. This we could also already see in the
boxplot (large boxes).

```{r density_ozone_per_month}
# Load the function
source("../R/airquality/function.density.plot.R")

# Function call
density.airquality(airquality_mutated)
```

We detect many missing values. For a stringent analysis we need to know
which values are missing. Are there a systematic error in the instrument?
Or simple forgot someone to write down the measurement? For that we use a
bar-plot which replace all NA in -1. Now we can easily see which values are
missing. We are missing a lot of values for June. To create new values
through linear interpolation is not an option because there is (1) to many
NA and (2) the gap between to values is to big. We would only falsify our
analysis.

```{r bar_plot_missing_ozone_per_month}
# Load the function
source("../R/airquality/function.na.identifier.R")

# Function call
na.identifier(airquality_mutated )
```

## Correlation

Now we have a impression of our data. Now we want to know whether the
variables of the hypothesis are promisingly. For this purpose we create a
correlation matrix. If we do not know if the ozone distribution is linear
or monotonic, we use the "Pearson" and the "Spearman" method. The graphic
shows us a high positiv correlation between ozone concentration vs.
temperature (R = 0.7, $\rho=0.77$) and a high negative ozone concentration
vs. wind (R = -0.6 and $\rho=-0.59$). There is also a positive correlation
between ozone concentration and solar radiation (R = 0.35 and $\rho=0.35$)
but not so strong. The difference between the methods are not negligible.
In this exercise, we would use the "Spearman" method because it describes
the correlation a little bit better. Nevertheless, the chosen variables are
promisingly.

```{r correlation_plot}

source("../R/airquality/function.correlation.plot.R")

correlation.plot(airquality_mutated, c("pearson"))

correlation.plot(airquality_mutated, c("spearman"))
```

## Regression models

Simple linear regression

```{r linear_regression_model_ozone_temperature}
# Load the function
source("../R/airquality/function.regression.calculater.R")

# Function call ozone vs. Date
regression.calculater(airquality_mutated, 4, 1)

# Function call ozone vs. Temperature
regression.calculater(airquality_mutated, 4, 5)

# Function call ozone vs. Wind
regression.calculater(airquality_mutated, 4, 6)

# Function call ozone vs. Solar radiation
regression.calculater(airquality_mutated, 4, 7)
```

For a better interpretation the model can be visualized:

```{r linear_regression_model_ozone_time}
# Load the function
source("../R/airquality/function.regression.R")

# Function calls
airquality.regression(airquality_mutated, Day, c("Day"))

airquality.regression(airquality_mutated, Temp, c("Temp [K]"))

airquality.regression(airquality_mutated, Wind, c("Wind [m/s]"))

airquality.regression(airquality_mutated, Solar.R, c("Solar.R [W/m^2]"))
```

## Results and Discussion

## Bibliography

Brönnimann Stefan (2013): Ozon in der Atmosphäre. Bern, Haupt Verlag. DOI:
10.480/GB2013.03
