---
title: "Report_Exercise_2"
author: "Patrick Bigler"
date: "2023-04-10"
output: html_document
editor_options: 
  markdown: 
    wrap: 75
---

# Report Exercise: "Data visualization"

Course: Applied Geo-data Science at University of Bern (Institute of
Geography)

Supervisor: Prof. Dr. Benjamin Stocker

Adviser: Dr. Koen Hufkens, Pepa Aran, Pascal Schneider

[You have questions to the workflow? Contact the Author:]{.underline}

Author: Bigler Patrick (patrick.bigler1\@students.unibe.ch)

Matriculation number: 20-100-178

Reference: Report Exercise 2 (Chapter 4)

## Introduction and Hypothesis

Air quality depends on factors such as particulate matter (PM10 or PM2.5),
nitrogen dioxide $(NO_2)$ or ozone $(O_3)$. In this exercise, only the
tropospheric ozone is considered for air quality. Ozone is not a health
hazard per se but a high concentration over a longer period of time can
affect the health (Brönnimann 2013, 23). But what does that mean? For
Switzerland, the threshold value is $120\: \frac{\mu g}{m^3}$. This value
is the average for an hour and can be exceeded only once per year (it
exists other definitions for a threshold, but here we use this definition
(Brönnimann 2013, 23). However, this unit depends on the air pressure,
which decreases with altitude. The volume mixing ratio with the unit parts
per million is more suitable which is about 60 ppb which we use in this
exercise. (Brönnimann 2013, 26).

Ozone is a molecule consisting of 3 oxygen atoms and has various physical
(absorption of UV and IR radiation) and chemical properties (high
reactivity). Ozone emerges due to a 2 phase chemical reaction (Brönnimann
2013, 75):

1.  $NO_2 + h\nu \left(\lambda < 420 nm \right) \rightleftharpoons NO + O$

2.  $NO + O_3 \rightleftharpoons NO_2 + O_2$

From the reaction equations it follows that UV radiation, oxygen and
nitrogen dioxide are needed for the formation of ozone . The first two are
available in almost unlimited quantities. Thus, the production depends
mainly on nitrogen dioxide. For the planetary boundary layer (and therefore
for the surface as well), the motorized traffic is the main source
(Brönnimann 2013, 95). Lastly, many chemical reactions are temperature
dependent. This is also true for ozone. The higher the temperature, the
more the balance of the reaction shifts towards ozone production and the
concentration increases (Brönnimann 2013, 95).

In summary, the ozone concentration depends probably on temperature (shift
of the balance), solar radiation (mostly UV-radiation) and the
concentration of nitrogen dioxide (which depends on traffic and wind). With
these considerations, a data set on air-quality can be studied with a focus
on ozone.

This exercise is supposed to answer the following research question:

*"Quantify the dependency of ozone concentration from temperature, solar
radiation and wind"*

## Methods

The method is quite straigth forward.

1.  We make some quality control calls
2.  We check the source because the data set is a second source

First, there is a little bit of data quality control. We have to know, what
variables contains the data set and how many missing values are there.
After that we check the sources and decide if we want to use the data or if
there any changes necessary. Then we make some statistic calls and extract
and visualize the main statistic parameters. With this knowledge we will be
able to describe the ozone with selected variables.

## Programming and data evaluation

### Packages

The following code chunk contains all packages we need. Important is the
package "conflicted". It enables us to chose if different functions have
the same call but do not make the same thing (a function in a certain
package can have the same name as a function from another package).

```{r activate_the_packages_which_are_required}
source("../R/general/packages.R")
```

### Original source of the file

The data is example data frame in the R package "datasets". First, the data
frame is a secondary source an we must know about the contain, the values,
the units and, of course, the primary sources. For that we can use R to
gather all these information. With help() we see that the varibales do not
have the SI-units. We also check the NAs. With str() we can see that there
are NAs. We call a function to visualize the NAs.

```{r shows_the_data_frame_airquality}
# Read the data
airquality <- datasets::airquality

# Ask for details (units, source and references)
help("airquality")

# check data type of every variable in data frame
str(airquality)

# Load the function
source("../R/general/function.visualize.na.values.R")

# Function call
visualize.na.values(airquality)
```

The data frame shows different variables at different times. Start was
05.01.1973 (a Tuesday) and ends at 09.30.1973 and each day has an entry (it
could be "NA" as well). We distinguish the variable according to the source
and describing them.

Data for ozone were collected from the New York State Department of
Conservation.

-   Ozone [ppb] represent the average of the concentration between 1pm -
    3pm at Roosevelt Island, NYC.

Data for the meteorological data were collected from the US. National
Weather Service.

-   Wind [mph] is calculated as the average wind speed at 7am and 10am at
    La Guardia Airport, NYC

-   Temperature [degrees Fahrenheit] is the maximum daily temperature at la
    Guardia Airport, NYC

-   Solar. R [Langley] average in the frequency band 4000-7700 Angstroms
    (wavelength: 400-770 nano meter) from 8am to noon in Central Park NYC.

### Editing the file in SI-Units

It is strongly recommended to use the SI unit system because it is
simplifies the reproducibility and the interpretation and discussion of the
data. It follows the formulas to calculates the values from there units
into the SI-units system

-   For temperature the following formula was used:
    $$\frac{x \left[°Fahrenheit \right]- 32}{1.8} +273.15 \Longleftrightarrow x\left[Kelvin \right]$$

-   For wind speed the following formula was used :
    $$1 \left[\frac{mile}{hour} \right] \Longleftrightarrow 0.44704 \left[\frac{meter}{second}\right]$$

-   For solar radiation the following formula was used :
    $$1 \left[ Langley \right] \Longleftrightarrow 41840 \left[ \frac{Joule}{m^2}\right]$$

```{r editing_the_data_frame_in_SI_units}
# Load the function
source("../R/airquality/function.unit.converter.R")

# Function call
airquality_mutated <- unit.converter(airquality)
```

## Statistics

Now that all units have been converted to SI units, the question about the
position, scattering, and distribution parameters arises. For the ozone
concentration is the median the right position parameter because it is
robust against outliers. This is important because a short time higher
concentration of ozone is probably not a problem for the human well-being
but for a long term it could be. If we use the mean, we loss the overall
picture due to a big one time outlier. If the median has been used it
should be used the IQR as a scattering parameter. But we are also
interested in the maximal values. A environment with a very high ozone
concentration can be toxic (even a short time exposure).

The data shows us that the monthly median (may = 18 ppb, Jun = 23 ppb, Jul
= 60 ppb, Aug = 52 ppb, Sept = 23 ppb) is lower or equal to the threshold
value (60 ppm). Therefore, there the air-quality is acceptable. But there
are measurements in every month which exceed the threshold value (.75
quantile, maximum ). That means that there are isolated days when the ozone
concentration is problematic. If we compare the span (167 ppb) with the IQR
(45.25 ppb) we notice that the IQR is clearly smaller than the span. That
means that we indeed have outliers but the the data core is small.

```{r statistical_parameters}
# Monthly median for temperature, ozone, solar and wind
airquality_mutated |>
# Calculate the median for the ozone concentraion  
  summarise(median.temp = median(Temp, na.rm = TRUE),
            median.wind = median(Wind, na.rm = TRUE),
            median.ozone = median(Ozone, na.rm = TRUE),
            Min.ozone = min(Ozone, na.rm = TRUE),
            Q1.ozone = quantile(Ozone, 0.25, na.rm = TRUE),
            Q2.ozone = quantile(Ozone, 0.5, na.rm = TRUE),
            Q3.ozone = quantile(Ozone, 0.75, na.rm = TRUE),
            Max.ozone = max(Ozone, na.rm = TRUE),
            IQR.ozone = IQR(Ozone, na.rm = TRUE),
            Span = max(Ozone, na.rm = TRUE) - min(Ozone, na.rm = TRUE),
            Average.ozone = mean(Ozone, na.rm = TRUE))
```

We can see the same in boxplots. But in a box-whisker plot are some
information hidden because its shows only the percentiles. Therefore, a
jitters plot is added. The plot shows us that there are outliers (red
triangle), but the boxes for May, June and September are small which means
the data scattering is small. The situation for July and August is
different. The large box indicates a large IQR for these months and the
data scattering is relatively high. Additional, the plot shows also the
mean value. It is higher than the median for each box. Therefore, the
skewness is for all months positive. Please note that there are only 9
measurements available for June.

```{r boxplot_ozone_per_month}
# Load the function
source("../R/airquality/function.boxplot.R")

# Function call
airquality.boxplot(airquality_mutated)
```

The density plot shows us that the data have a big span and a positive
skewness. But the core of the data (area with a high density) has an
acceptable span (which is basically the IQR). The graphs shows us also that
the density for the months May, June and September is bi-modal. That means,
that there are outliers (which we already see in the boxplot). July and
August are mono-modal but have a large span. This we could also already see
in the boxplot (large boxes).

```{r density_ozone_per_month}
# Load the function
source("../R/airquality/function.density.plot.R")

# Function call
density.airquality(airquality_mutated)
```

We detect many missing values. For a stringent analysis we need to know
which values are missing and where they are. Are there a systematic error
in the instrument? Or simple forgot someone to write down the measurement?
For that we use a bar-plot which replace all NA in -1. Now we can easily
see which values are missing. We are missing a lot of values for June. To
create new values through linear interpolation is not an option because
there is (1) to many NA and (2) the gap between to values is to big. We
would only falsify our analysis.

```{r bar_plot_missing_ozone_per_month}
# Load the function
source("../R/airquality/function.na.identifier.R")

# Function call
na.identifier(airquality_mutated )
```

## Correlation

Now we have a impression of our data. Now we want to know whether the
variables of the hypothesis are promisingly. For this purpose we create a
correlation matrix. Because we do not know if the ozone distribution is
linear or monotonic, we use the "Pearson" and the "Spearman" method. The
graphic shows us a high positive correlation between ozone concentration
vs. temperature (r = 0.7, $\rho=0.77$) and a high negative ozone
concentration vs. wind (r = -0.6 and $\rho=-0.59$). There is also a
positive correlation between ozone concentration and solar radiation (R =
0.35 and $\rho=0.35$) but not so strong. The difference between the methods
are not negligible. In this exercise, we would use the "Spearman" method
because it describes the correlation a little bit better. Nevertheless, the
chosen variables are promisingly.

```{r correlation_plot}

source("../R/airquality/function.correlation.plot.R")

correlation.plot(airquality_mutated, c("pearson"))

correlation.plot(airquality_mutated, c("spearman"))
```

## Regression models

We want to be able to predict new values. For that we need a model. Here we
create a simple linear regression model. For that we calculate four
different models.

The first model is ozone vs. time. The model explains only about 2.5 % of
the variance (RSQ = 0.024). Both the coefficient and the intercept are not
significant (p-value \> 0.05) and we can not reject H0 (Is the p-value
insignificant, then it indicates there is insufficient evidence in the
sample to conclude that a non-zero correlation exists). We reject that
model.

The second model is ozone vs. temperature. The model explains about 48% of
the variance. That is not so bad but far from perfect. The p-value is \<
0.05 and therefore significant (even on a 99.9 % level). We can reject H0.
We accept the model temporarily.

The third model is ozone vs. wind. The RSQ is about 0.36. The model
explains about 36 % of the variance. The p-value is \< 0.05 and the
coefficient and the intercept are significant. We reject H0 and accept the
model temporarily.

The last model is ozone vs. solar radiation. This model is not as good as
the second and the third model. RSQ is only about 0.12 what explains only
12 % of the variance. The p-value is \< 0.05 but the intercept is only on a
99 % level significant (actually this is a good value but it is lower
compared to the second an third model). We reject H0 but the model is not
so good so we do not use it.

```{r linear_regression_model_ozone_temperature}
# Load the function
source("../R/airquality/function.regression.calculater.R")

# Function call ozone vs. Date
regression.calculater(airquality_mutated, 4, 1)

# Function call ozone vs. Temperature
regression.calculater(airquality_mutated, 4, 5)

# Function call ozone vs. Wind
regression.calculater(airquality_mutated, 4, 6)

# Function call ozone vs. Solar radiation
regression.calculater(airquality_mutated, 4, 7)
```

For a better interpretation the model can be visualized.

```{r linear_regression_model_ozone_time}
# Load the function
source("../R/airquality/function.regression.R")

# Function calls
airquality.regression(airquality_mutated, Day, c("Day"))

airquality.regression(airquality_mutated, Temp, c("Temp [K]"))

airquality.regression(airquality_mutated, Wind, c("Wind [m/s]"))

airquality.regression(airquality_mutated, Solar.R, c("Solar.R [W/m^2]"))
```

## Discussion

In this short analysis we can see that ozone probably depends on
temperature, wind and solar radiation. Although the RSQ and the p-value was
not bad for the second and third model, none of the models is convincing
enough to use as an explanatory model. We need better approaches. One
possibility would be multiple regression or machine learning approaches.
However, these are not used here because they will be introduced later.

Although the RSQ for the second and third model was not so bad, the
visualization shows a new problem. The RSQ in the model is valid for the
whole time-period. However, in the visualization you can see that the RSQ
is highly fluctuating. It could be a good idea, when we chose a new
approach for each month. If we analyse more data (maybe over 20-50 years)
we could discover pattern. May be we must use for the month May another
variable than for June. With only this data it is impossible to say what we
should to. If we look for patterns in this data, any pattern found would
probably be random.

But we can take a look on this specific data set and analyse the year 1973.
First we must discuss the source. It is problematic that not all variables
were measured at the same location. The ozone was measured on Roosvelt
Island. It is an island on the East River in NYC. But the temperature and
wind were measured at La Guardia Airfield. The solar irradiance in Central
Park. The measuring stations are up to 10 km apart and this is in the
middle of a highly city. There is no guarantee that the meteorological
conditions are the same. There is probably more wind at the airport than at
Roosvelt Island. It would be better to collect new, up-to-date data
ourselves and measure all variables at the same place. It would be even
better if we implemented an urban measurement network. This would probably
improve the model massively.

Second we can see, that the median is lower than the threshold for all
months. The temperature is a important factor for the ozone concentration.
The same is true for the wind. In the correlation plot showed us that there
is a correlation between temperature and wind (what makes sense in a
meteorological way). We also know, that the ozone concentration do not
correlate with time. This makes also sense. Although the ozone
concentration is higher during summer (summer smog), there is not relation
to the day of a month. The concentrations reacts on physical parameter like
radiation, wind and temperature. It makes absolutely sense, that the
temperature is higher Jun, Jul, Aug and Sept than May. The ozone
concentration is tends to be higher if the temperature is high. In a urban
space this effect will be amplified due to heat-islands. A city like NYC is
predestined for such an effect.

As a conclusion we can say that the air-quality is less bad as expected.
But there are days, where the concentration is to high (true for all
months). The city of NY should take measures to prevent heat islands. This
could lower temeprature, which inhibits ozone production, thus lowering
concentration. Further, the source material should be limited. With less
traffic, NOX could be lowered, inhibiting ozone production. But first more
data should be collected. We therefore recommend to build up a city-wide
monitoring network, which should include more air quality parameters, such
as particulate matter. This would give us an overall picture and allow us
to take evidence-based measures.

## Bibliography

Brönnimann Stefan (2013): Ozon in der Atmosphäre. Bern, Haupt Verlag. DOI:
10.480/GB2013.03
