---
title: "Report_Exercise_2"
author: "Patrick Bigler"
date: "2023-04-10"
output: html_document
editor_options: 
  markdown: 
    wrap: 75
---

# Report Exercise to Chapter 4

Author: Patrick Bigler

Martikelnummer: 20-100-178

## Introduction and Hypothesis

Air quality depends on factors such as particulate matter (PM10 or PM2.5),
nitrogen dioxide $(NO_2)$ or ozone $(O_3)$. In this exercise, only the
tropospheric ozone is considered for air quality. Ozone is not a health
hazard per se but a high concentration over a longer period of time can
affect the health. (Brönnimann 2013, 23). But what does that mean? For
Switzerland, the threshold value is $120\: \frac{\mu g}{m^3}$ over eight
hours (Brönnimann 2013, 23). However, this unit depends on the air
pressure, which decreases with altitude. The volume mixing ratio with the
unit parts per million is more suitable which is about 60 ppb which we use
in this exercise. (Brönnimann 2013, 26).

Ozone is a molecule consisting of 3 oxygen atoms and has various physical
(absorption of UV and IR radiation) and chemical properties (high
reactivity). Ozone emerges due to a 2 phase chemical reaction (Brönnimann
2013, 75):

1.  $NO_2 + h\nu \left(\lambda < 420 nm \right) \rightleftharpoons NO + O$

2.  $NO + O_3 \rightleftharpoons NO_2 + O_2$

From the reaction equations it follows that UV radiation, oxygen and
nitrogen dioxide are needed for the formation of ozone . The first two are
available in almost unlimited quantities. Thus, the production depends
mainly on nitrogen dioxide. For the planetary boundary layer (and therefore
for the surface as well), it is motorized traffic the main source.
(Brönnimann 2013, 95). Lastly, many chemical reactions are temperature
dependent. This is also true for ozone. The higher the temperature, the
more the balance of the reaction shifts towards ozone production and the
concentration increases (Brönnimann 2013, 95).

In summary, the ozone concentration depends probably on temperature (shift
of the balance), solar radiation (mostly IR-radiation) and the
concentration of nitrogen dioxide (which depends on traffic and wind). With
these considerations, a data set on air-quality can be studied with a focus
on ozone.

This exercise is supposed to answer the following research question:

*"Quantify the dependency of ozone concentration from temperature, solar
radiation and wind"*

## Packages

The following code chunk contains all packages we need for this exercise.
Important is the package "conflicted". It shows us if the same function
name is used in another package for another function. If there are such a
conflict, we can (1) call the package name before the function call or (2)
we can set a priority. If there any conflict in this exercise, we will set
a preference in the first line of the code chunk.

```{r activate_the_packages_which_are_required}
source("../R/packages.R")
```

## Original source of the file

The data is example data frame in the R package "datasets". First, the data
frame is a secondary source an we must know about the contain, the values,
the units and, of course, the primary sources. For that we can use R to
gather all these information.

```{r shows_the_data_frame_airquality}
# Read the data
airquality <- datasets::airquality

# Ask for details (units, source and references)
help("airquality")

# Check weather there are missing values
is.na(airquality)

# check data type of every variable in data frame
str(airquality)
```

The data frame shows different variables at different times. Start was
05.01.1973 (a Tuesday) and ends at 09.30.1973 and each day has an entry (it
could be "NA" as well). We distinguish the variable according to the source
and describing them.

Data for ozone were collected from the New York State Department of
Conservation.

-   Ozone [ppb] represent the average of the concentration between 1pm -
    3pm at Roosevelt Island, NYC.

Data for the meteorological data were collected from the US. National
Weather Service.

-   Wind [mph] is calculated as the average wind speed at 7am and 10am at
    La Guardia Airport, NYC

-   Temperature [degrees Fahrenheit] is the maximum daily temperature at la
    Guardia Airport, NYC

-   Solar. R [Langley] average in the frequency band 4000-7700 Angstroms
    (wavelength: 400-770 nano meter) from 8am to noon in Central Park NYC.

## Editing the file in SI-Units

It is strongly recommended to use the SI unit system because it is
simplifies the reproducibility and the interpretation and discussion of the
data. It follows the formulas to calculates the values from there units
into the SI-units system

-   For temperature the following formula was used:
    $$\frac{x \left[°Fahrenheit \right]- 32}{1.8} +273.15 \Longleftrightarrow x\left[Kelvin \right]$$

-   For wind speed the following formula was used :
    $$1 \left[\frac{mile}{hour} \right] \Longleftrightarrow 0.44704 \left[\frac{meter}{second}\right]$$

-   For solar radiation the following formula was used :
    $$1 \left[ Langley \right] \Longleftrightarrow 41840 \left[ \frac{Joule}{m^2}\right]$$

```{r editing_the_data_frame_in_SI_units}

airquality_mutated <- airquality |>
# Change the units and make the data tidy 
  mutate(Group = Month,
         Temp = (Temp - 32) / 1.8 + 273.15, 
         Wind = Wind * 0.44704,
         Solar.R = Solar.R * 41840,
         Date = make_date(1973, Group, Day)) |>
  select("Date", "Group", "Ozone", "Temp", "Wind", "Solar.R")

# Define a order for the rows. We need it for the x-axis of the graphs
airquality_mutated$Group <- factor(airquality_mutated$Group, 
                                   levels =  c(5, 6, 7, 8, 9), 
                                   labels = c("May", "Jun", "Jul", "Aug", "Sept"))
```

## Statistics

Now that all units have been converted to SI units and the work with the
variables is legitimate, the question about the position, scattering and
distribution parameters arises. For the ozone concentration is the median
the right position parameter because it is robust against outliers. This is
important because a short time higher concentration of ozone is probably
not a problem for the human well-being but for a long term it could be. If
we use the mean, we loss the overall picture due to a big one time outlier.
If the median has been used it should be used the IQR as a scattering
parameter.

It shows us that the monthly median (may = 18 ppb, Jun = 23 ppb, Jul = 60
ppb, Aug = 52 ppb, Sept = 23 ppb) is lower or equal to the threshold value
(60 ppm). Therefore, there the air-quality is acceptable. But there are
measurements in every month which exceed the threshold value (.75 quantile,
maximum ). That means that there are isolated days when the ozone
concentration is problematic. If we compare the span (167 ppb) with the IQR
(45.25 ppb) we notice that the IQR is clearly smaller than the span . That
means that we indeed have outliers but the the data core is small.
Therefore, the data quality is acceptable.

```{r statistical_parameters}
#monthly median for temperature, ozone, solar and wind
airquality_mutated |>
#calculate the median for the ozone concentraion  
  summarise(median.temp = median(Temp, na.rm = TRUE),
            median.wind = median(Wind, na.rm = TRUE),
            median.ozone = median(Ozone, na.rm = TRUE),
            Min.ozone = min(Ozone, na.rm = TRUE),
            Q1.ozone = quantile(Ozone, 0.25, na.rm = TRUE),
            Q2.ozone = quantile(Ozone, 0.5, na.rm = TRUE),
            Q3.ozone = quantile(Ozone, 0.75, na.rm = TRUE),
            Max.ozone = max(Ozone, na.rm = TRUE),
            IQR.ozone = IQR(Ozone, na.rm = TRUE),
            Span = max(Ozone, na.rm = TRUE) - min(Ozone, na.rm = TRUE),
            Average.ozone = mean(Ozone, na.rm = TRUE))
```

We can see this also in the boxplot. But in a box-whisker plot are some
information hidden because its shows only the percentiles. Therefore, a
jitters plot is added. The plot shows us that there are outliers, but the
boxes for May, June and September are small which means the data scattering
is small. The situation for July and August is different. The large box
indicates a large IQR for these months and the data scattering is
relatively high. Additional, the plot shows also the mean value. It is
higher than the median for each box. Therefore, the skewness is for all
months positive.

```{r boxplot_ozone_per_month}
# use conflicted package to prefer a package
conflicts_prefer(dplyr::filter)
conflicts_prefer(ggplot2::annotate)

# use ggplot to visualise the distribution of ozone for each month
ggplot(data = airquality_mutated, aes(x = factor(Group), y = Ozone)) +
# Boxplot with an individual setup in transparency, size and width
  geom_boxplot(fill = "skyblue", alpha = 0.5, lwd = 0.3, width = 0.5,
# red colour, individual shape and size for outliers               
               outlier.color = "red", outlier.shape = 17, outlier.size = 2) +
# Set new limits for the y-axis (we want it a little bit higher than max. value)
  scale_y_continuous(limits = c(0, 200)) +
# count and write how many values are in the boxplot (without NA!)  
  annotate("text",
           x = 1:length(table(airquality_mutated$Group)),
           y = 180,
           label = c(paste("n =", nrow(airquality_mutated |>
         drop_na(Ozone) |>
         filter(Group == "May") |>
         select(Ozone))) , paste("n =", nrow(airquality_mutated |>
         drop_na(Ozone) |>
         filter(Group == "Jun") |>
         select(Ozone))), paste("n =", nrow(airquality_mutated |>
         drop_na(Ozone) |>
         filter(Group == "Jul") |>
         select(Ozone))), paste("n =", nrow(airquality_mutated |>
         drop_na(Ozone) |>
         filter(Group == "Aug") |>
         select(Ozone))), paste("n =", nrow(airquality_mutated |>
         drop_na(Ozone) |>
         filter(Group == "Sept") |>
         select(Ozone)))), col = "black", vjust = -1) +
# add an errorbar and make a individual setup 
  stat_boxplot(geom = "errorbar", size = 0.3, width = 0.3) +
# add mean and make a individual setup  
  stat_summary(fun.y = mean, geom = "point", shape = 23, size = 4, color = "orange") +
# make the values almost transparent and force the values with width statemant in a colomn
  geom_jitter(width = 0, alpha = 0.1) +
# create and add a fine dotdashed line for the threshold values for Switzerland as a reference
  geom_hline(yintercept = 60, color = "darkmagenta", size = 0.3, linetype = "dotdash") +
# create and add a dotted line for the median for the ozone concentration 
  geom_hline(yintercept =  median(airquality_mutated$Ozone, na.rm = TRUE), 
             color = "royalblue", size = 0.3, linetype = "dotdash") +
# add label for the Threshold, hjust for a left side alignment
  annotate("text", x =  0.5 , y = 170, hjust = 0, 
           label = "Threshold value [60 ppb]", color = "darkmagenta") +
# add label for the median, hjust for a left side alignment
  annotate("text", x = 0.5 , y =  155, hjust = 0, 
           label = "Overall median [31.5 ppm]", color = "royalblue") +
# add label for the outliers, hjust for a left side alignment
  annotate("text", x = 0.5 , y =  140, hjust = 0, 
           label = "Outliers", color = "red") +
# add label for the mean, hjust for a left side alignment
  annotate("text", x = 0.5 , y =  125, hjust = 0, 
           label = "Mean", color = "orange") +
# label the graphic and axes
  labs(title = "Ozone concentration per month",
       subtitle = "At Roosevelt Island, NYC [Average concentration between 1pm and 3pm]",
       x = "Time [1973]", y = "Ozone concentration [ppb]",
       caption = "AGDS Report Exercise 2 (Chapter 4)") +
# chose a background for the graph 
  theme_bw() +
# write the title bold  
  theme(plot.title = element_text(size = 15, face = "bold")) +
# add panel boarder for the graph 
  theme(panel.border = element_rect(colour = "black", fill = NA, linewidth = 1)) +
# add a individual panel boarder around the graph  
  theme(plot.margin = margin(0.3, 1, 0.3, 0.3, "cm"),
  panel.background = element_rect(fill = "white"),
  plot.background = element_rect(fill = "grey90",colour = "black", linewidth = 1)) 
```

The density plot shows us that the data have a big span. But the core of
the data (area with a high density) has an acceptable span (which is
basically the IQR). The graphs shows us also that the density for the
months May, June and September is bi-modal. That means, that there are
outliers (which we already see in the boxplot). July and August are
mono-modal but have a large span. This we could also already see in the
boxplot (large boxes).

```{r density_ozone_per_month}
airquality_mutated|>
ggplot(aes(x = Ozone, group = Group, fill = Group)) +
# add the geom_density and create an individual setup
    geom_density(adjust = 1.5, alpha = 0.4) +
# add title and subtitle and label the axes  
    labs(x = "Ozon [ppb]", y = "Density", 
    title = "Density of the ozone concentration",
    subtitle = "At Roosevelt Island, NYC [Average concentration between 1pm and 3pm]",
    caption = "AGDS Report Exercise 2 (Chapter 4)") +
# add a vline with an individual setup  
    geom_vline(xintercept = 60, size = 0.3, 
               linetype = "dotdash", color = "darkmagenta") +
    geom_vline(xintercept = median(airquality_mutated$Ozone, na.rm = TRUE), 
               size = 0.3, linetype = "dotdash", color = "royalblue") +
    ggplot2::annotate("text", x = 110 , y =  0.0218, hjust = 0, 
           label = "Overall median [31.5 ppb]", color = "royalblue") +
    ggplot2::annotate("text", x = 110 , y =  0.0235, hjust = 0, 
           label = "Threshold value [60 ppb]", color = "darkmagenta") +
# chose a background and add a panel boarder  
    theme_bw() +
# plot the title bold 
    theme(plot.title = element_text(size = 15, face = "bold")) +
# add a panel boarder  
    theme(panel.border = element_rect(colour = "black", fill = NA, linewidth = 1)) +
    theme( legend.background = element_rect( fill = "lightblue", 
                size = 0.5, linetype = "solid", colour = "black")) +
# add a panel boader around the whole graph  
    theme(plot.margin = margin(0.3, 0.3, 0.3, 0.3, "cm"),
    panel.background = element_rect(fill = "white"),
    plot.background = element_rect(fill = "grey90",colour = "black", linewidth = 1))   
```

We detect many missing values. For a stringent analysis we need to know
which values are missing. Are there a systematic error in the instrument?
Or simple forgot someone to write down the measurement? For that we use a
bar-plot which replace all NA in -1. Now we can easily see which values are
missing. We are missing a lot of values for June. To create new values
through linear interpolation is not an option because there is (1) to many
NA and (2) the gap between to values is to big. We would only falsify our
analysis.

```{r bar_plot_missing_ozone_per_month}
# create a bar plot and mark the missing values with -1
airquality_mutated |>
  ggplot(aes(x = Date, y = Ozone, group = Group, color = Group)) +
# create a geom bar  
  geom_bar(stat = "identity") +
# make a geom with points  
  geom_point(size = 1.5) +
# make a geom with the missing values. change NA to -1  
  geom_point(data = airquality_mutated |>
  mutate_all( ~replace_na(.,-1)) |>
  dplyr::filter(`Ozone` == -1) , size = 1.5, color = "red") +
  labs(x = "Time [1973]", y = "Ozon [ppb]", 
  title = "Daily ozone concentration",
  subtitle = "At Roosevelt Island, NYC [daily average concentration between 1pm and 3pm]",
  caption = "AGDS Report Exercise 2 (Chapter 4)") +
# chose a background and add a panel boarder  
  theme_bw() +
# plot title bold  
  theme(plot.title = element_text(size = 15, face = "bold")) +
# add an individual panel boarder  
  theme(panel.border = element_rect(colour = "black", fill = NA, linewidth = 1)) +
  theme( legend.background = element_rect( fill = "lightblue", 
                size = 0.5, linetype = "solid", colour = "black")) +
# add an individual panel boarder around the whole graph  
  theme(plot.margin = margin(0.3, 0.3, 0.3, 0.3, "cm"),
  panel.background = element_rect(fill = "white"),
  plot.background = element_rect(fill = "grey90",colour = "black", linewidth = 1))
```

## Correlation

Now we have a impression of our data. Now we want to know whether the
variables of the hypothesis are promisingly. For this purpose we create a
correlation matrix. If we do not know if the ozone distribution is linear
or monotonic, we use the "Pearson" and the "Spearman" method. The graphic
shows us a high positiv correlation between ozone concentration vs.
temperature (R = 0.7, $\rho=0.77$) and a high negative ozone concentration
vs. wind (R = -0.6 and $\rho=-0.59$). There is also a positive correlation
between ozone concentration and solar radiation (R = 0.35 and $\rho=0.35$)
but not so strong. The difference between the methods are not negligible.
In this exercise, we would use the "Spearman" method because it describes
the correlation a little bit better. Nevertheless, the chosen variables are
promisingly.

```{r correlation_plot}
# correlation: Method "pearson"
C1 <- airquality_mutated |>
  select( Ozone, Wind, Temp, Solar.R) |>
# chose the right method  
  cor(use = "pairwise", method = "pearson") |>
# plot lower triangle with the correlation coefficients
  ggcorrplot(hc.order = TRUE,
           type = "lower", lab = TRUE) +
# label the graphic
  labs(title = "Correlation overview", 
       subtitle = "Method: Pearson",
       caption = "AGDS Report Exercise 2 (Chapter 4)") +
# plot the title bold
  theme(plot.title = element_text(size = 15, face = "bold")) +
# add an individual panel boarder  
  theme(panel.border = element_rect(colour = "black", fill = NA, linewidth = 1)) +
  theme( legend.background = element_rect( fill = "lightblue", 
                size = 0.5, linetype = "solid", colour = "black")) +
# add an individual panel boarder around the whole graph
  theme(plot.margin = margin(0.3, 0.3, 0.3, 0.3, "cm"),
  panel.background = element_rect(fill = "white"),
  plot.background = element_rect(fill = "grey90",colour = "black", linewidth = 1))  

# correlation: Method "spearman"
C2 <- airquality_mutated |>
  select( Ozone, Wind, Temp, Solar.R) |>
# chose the right method  
  cor(use = "pairwise", method = "spearman") |>
# plot lower triangle with the correlation coefficients
  ggcorrplot(hc.order = TRUE,
           type = "lower", lab = TRUE) +
# label the graphic
  labs(title = "Correlation overview",
       subtitle = "Method: Spearman",
         caption = "AGDS Report Exercise 2 (Chapter 4)") +
# plot the title bold
  theme(plot.title = element_text(size = 15, face = "bold")) +
# add an individual panel boarder  
  theme(panel.border = element_rect(colour = "black", fill = NA, linewidth = 1)) +
  theme( legend.background = element_rect( fill = "lightblue", 
                size = 0.5, linetype = "solid", colour = "black")) +
# add an individual panel boarder around the whole graph
  theme(plot.margin = margin(0.3, 0.3, 0.3, 0.3, "cm"),
  panel.background = element_rect(fill = "white"),
  plot.background = element_rect(fill = "grey90",colour = "black", linewidth = 1))

plot_grid(C1, C2, labels = c("A","B"))
```

## Regression models

Simple linear regression

```{r linear_regression_model_ozone_temperature}
#calculate a simple linear regression: ozone vs. temperature
model.fit.ozone.temp <- 
  lm(airquality_mutated$Ozone ~ airquality_mutated$Temp,
  airquality_mutated = airquality_mutated)
summary(model.fit.ozone.temp)

#calculate a simple linear regression: ozone vs. wind
model.fit.ozone.wind <- 
  lm(airquality_mutated$Ozone ~ airquality_mutated$Wind, 
  airquality_mutated = airquality_mutated)
summary(model.fit.ozone.wind)

#calculate a simple linear regression: ozone vs. solar radiation
model.fit.ozone.solar <- 
  lm(airquality_mutated$Ozone ~ airquality_mutated$Solar.R, 
  airquality_mutated = airquality_mutated)
summary(model.fit.ozone.solar)
```

For a better interpretation the model can be visualized:

```{r linear_regression_model_ozone_time}
P1 <- airquality_mutated |>
  ggplot(aes(x = Day, y = Ozone, color = Month)) +
  geom_point(alpha = 0.5) +
  scale_y_continuous(limits = c(0, 250)) +
#use smooth for a linear regression, se for standard error  
  geom_smooth(formula = y ~ x, method = "lm", se = FALSE) +
  stat_poly_eq(use_label(c("eq", "R2"))) +
  geom_hline(yintercept = 60, linetype = "dotdash", 
             size = 0.3, color = "darkmagenta") +
#lab the axes and giv the graphic a title  
  labs(title = "Ozone vs. Time",
       x = "Time [Day]", y = "Ozone [ppb]",
       caption = "AGDS Report Exercise 2 (Chapter 4)") +
  theme_bw() +
  theme(plot.title = element_text(size = 15, face = "bold")) +
  theme(panel.border = element_rect(colour = "black", 
       fill = NA, linewidth = 1)) +
  theme( legend.background = element_rect( fill = "lightblue", 
       size = 0.5, linetype = "solid", colour = "black")) +
  theme(plot.margin = margin(0.3, 0.3, 0.3, 0.3, "cm"),
  panel.background = element_rect(fill = "white"),
  plot.background = element_rect(fill = "grey90",colour = "black", linewidth = 1))   
```

```{r visualize_linear_regression_model_ozone_temperature}
P2 <- airquality_mutated |>
  ggplot(aes(x = Temp, y = Ozone, color = Month)) +
  geom_point(alpha = 0.5) +
  scale_y_continuous(limits = c(0, 250)) +
#use smooth for a linear regression, se for standard error  
  geom_smooth(formula = y ~ x, method = "lm", se = FALSE) +
  stat_poly_eq(use_label(c("eq", "R2"))) +
  geom_hline(yintercept = 60, linetype = "dotdash", 
          size = 0.3, color = "darkmagenta") +
#lab the axes and giv the graphic a title  
  labs(title = "Ozone vs. Temperature", 
          x = "Temperature [Kelvin]", y = "Ozone [ppb]",
       caption = "AGDS Report Exercise 2 (Chapter 4)") +
  theme_bw() +
  theme(plot.title = element_text(size = 15, face = "bold")) +
  theme(panel.border = element_rect(colour = "black", 
          fill = NA, linewidth = 1)) +
  theme( legend.background = element_rect( fill = "lightblue", 
          size = 0.5, linetype = "solid", colour = "black")) +
  theme(plot.margin = margin(0.3, 0.3, 0.3, 0.3, "cm"),
  panel.background = element_rect(fill = "white"),
  plot.background = element_rect(fill = "grey90",colour = "black", linewidth = 1))   
```

```{r visualize_linear_regression_model_ozone_wind}
P3 <- airquality_mutated |>
  ggplot(aes(x = Wind, y = Ozone, color = Month)) +
  geom_point(alpha = 0.5) +
  scale_y_continuous(limits = c(0, 250)) +
#use smooth for a linear regression, se for standard error  
  geom_smooth(formula = y ~ x, method = "lm", se = FALSE) +
  stat_poly_eq(use_label(c("eq", "R2"))) +
  geom_hline(yintercept = 60, linetype = "dotdash", 
        size = 0.3, color = "darkmagenta") +
#lab the axes and giv the graphic a title  
  labs(title = "Ozone vs. Wind",
       x = "Wind [m/s]", y = "Ozone [ppb]",
       caption = "AGDS Report Exercise 2 (Chapter 4)") +
  theme_bw() +
  theme(plot.title = element_text(size = 15, face = "bold")) +  
  theme(panel.border = element_rect(colour = "black", 
        fill = NA, linewidth = 1)) +
  theme( legend.background = element_rect( fill = "lightblue", 
        size = 0.5, linetype = "solid", colour = "black")) +
  theme(plot.margin = margin(0.3, 0.3, 0.3, 0.3, "cm"),
  panel.background = element_rect(fill = "white"),
  plot.background = element_rect(fill = "grey90",colour = "black", linewidth = 1)) 
```

```{r visualize_regression_model_solar.r}
P4 <- airquality_mutated |>
  ggplot(aes(x = Solar.R, y = Ozone, color = Month)) +
  geom_point(alpha = 0.5) +
  scale_y_continuous(limits = c(0, 250)) +
#use smooth for a linear regression, se for standard error  
  geom_smooth(formula = y ~ x, method = "lm", se = FALSE) +
  stat_poly_eq(use_label(c("eq", "R2"))) +
  geom_hline(yintercept = 60, linetype = "dotdash", 
        size = 0.3, color = "darkmagenta") +
#lab the axes and giv the graphic a title  
  labs(title = "Ozone vs. solar irradiation", 
       subtitle = "Linear Regression Model",
       x = "solar irradiation [J/m^2]", y = "Ozone [ppb]",
       caption = "AGDS Report Exercise 2 (Chapter 4)") +
  theme_bw() +
  theme(plot.title = element_text(size = 15, face = "bold")) +  
  theme(panel.border = element_rect(colour = "black", 
        fill = NA, linewidth = 1)) +
  theme( legend.background = element_rect( fill = "lightblue", 
        size = 0.5, linetype = "solid", colour = "black")) +
  theme(plot.margin = margin(0.3, 0.3, 0.3, 0.3, "cm"),
  panel.background = element_rect(fill = "white"),
  plot.background = element_rect(fill = "grey90",colour = "black", linewidth = 1)) 
```

## Results and Discussion

## Bibliography

Brönnimann Stefan (2013): Ozon in der Atmosphäre. Bern, Haupt Verlag. DOI:
10.480/GB2013.03
