---
title: "Report Exercise: re_airquality"
author: "Patrick Bigler"
date: "2023-04-10"
output: 
  html_document:
    toc: yes
    toc_float: yes
    number_sections: yes
editor_options: 
  markdown: 
    wrap: 75
---

Course: Applied Geo-data Science at University of Bern (Institute of
Geography)

Supervisor: Prof. Dr. Benjamin Stocker

Adviser: Dr. Koen Hufkens, Pepa Aran, Pascal Schneider

Further information: <https://geco-bern.github.io/agds/>

[You have questions to the workflow? Contact the Author:]{.underline}

Author: Bigler Patrick (patrick.bigler1\@students.unibe.ch)

Matriculation number: 20-100-178

Reference: Report Exercise 2 (Chapter 4)

# Introduction

## Objectives

Visualizations are a proper way to communicate the results. The variety of
diagrams, colors and shapes is almost limitless. However, not all of them
are always equally suitable. It is essential to create the right graphic to
transport the message. With a realistic problem, the following goals should
be achieved:

-   Appropriate choice of visualisations for different data types and
    different aspects of the data

-   Grammar of graphics (ggplot2)

-   The proper use of colours in visualization

## Theory

Air quality depends on factors such as particulate matter (PM10 or PM2.5),
nitrogen dioxide $(NO_2)$ or ozone $(O_3)$. In this exercise, only the
tropospheric ozone is considered for air quality. Ozone is not a health
hazard per se but a high concentration over a longer period of time can
affect the health (Brönnimann 2013, 23). But what are high concentrations?
For Switzerland, the threshold value for ozone is
$120\: \frac{\mu g}{m^3}$. This value is the average for an hour and can be
exceeded only once per year. It exists other definitions for a threshold,
but here we use this definition (Brönnimann 2013, 23). However, this unit
depends on the air pressure, which decreases with altitude. The volume
mixing ratio with the unit parts per billion is more suitable which is
about 60 ppb (Brönnimann 2013, 26). Here, we use this value.

Ozone is a molecule consisting of 3 oxygen atoms and has various physical
(absorption of UV and IR radiation) and chemical properties (high
reactivity). Ozone emerges due to a 2 phase chemical reaction (Brönnimann
2013, 75):

1.  $NO_2 + h\nu \left(\lambda < 420 nm \right) \rightleftharpoons NO + O$

2.  $NO + O_3 \rightleftharpoons NO_2 + O_2$

From the reaction equations it follows that UV radiation, oxygen and
nitrogen dioxide are needed for the formation of ozone . The first two are
available in almost unlimited quantities. Thus, the production depends
mainly on nitrogen dioxide. For the planetary boundary layer (and therefore
for the surface as well), the motorized traffic is the main source
(Brönnimann 2013, 95). Lastly, many chemical reactions are temperature
dependent. This is also true for ozone. The higher the temperature, the
more the balance of the reaction shifts towards ozone production and the
concentration increases (Brönnimann 2013, 95).

In summary, the ozone concentration depends probably on temperature (shift
of the balance), solar radiation (mostly UV-radiation) and the
concentration of nitrogen dioxide (which depends on traffic and wind). With
these considerations, a data set on air quality can be studied with a focus
on ozone.

## Hypothesis

This exercise is supposed to answer the following research question:

*"Quantify the dependency of ozone concentration from temperature, solar
radiation and wind"*

# Methods

## Unit transformation into SI-Units

It is strongly recommended to use the SI unit system because it is
simplifies the reproducibility, the interpretation, and the discussion of
the data. It follows the formulas to transform the units into the SI-units
system.

-   For temperature the following formula was used:
    $$\frac{x \left[°Fahrenheit \right]- 32}{1.8} +273.15 \Longleftrightarrow x\left[Kelvin \right]$$

-   For wind speed the following formula was used :
    $$1 \left[\frac{mile}{hour} \right] \Longleftrightarrow 0.44704 \left[\frac{meter}{second}\right]$$

-   For solar radiation the following formula was used :
    $$1 \left[ Langley \right] \Longleftrightarrow 41840 \left[ \frac{Joule}{m^2}\right]$$

## R: Evaluation of the data

All analysis are done with the open source software R-Studio (Version
2022.12.0+353). This software use the language R to handle the data. For an
efficient processing of the data we use the R-package "Tidyverse (and
others).

# Programming and data evaluation

## Packages

The following code chunk contains all packages we need. Important is the
package "conflicted". It enables us to chose if different functions have
the same call but do not make the same thing (a function in a certain
package can have the same name as a function from another package).

```{r Packages, error=FALSE, message=FALSE, warning=FALSE}
source("../../R/general/packages.R")
```

## Read the file

First, we get the data and save it in our directory. But we must do this
only once. If the file exists, then we can read it directly. That is why we
implemented a if-else statement.

```{r Read_file, error=FALSE, message=FALSE, warning=FALSE}
# We want know, if a certain file already exists
name.of.file <- "../../data/re_airquality/airquality.csv"

# If do not exists such a file, create it!
if (!file.exists(name.of.file)){
  # Read the data
  airquality.raw <- datasets::airquality

  # Write a CSV file in the respiratory to save the data
  write_csv(airquality.raw, "../../data/re_airquality/airquality.csv")
  
  # Read the file
  airquality <- read_csv("../../data/re_airquality/airquality.csv")

  # If exists such a file, read it only!
  # Read the file
  }else{airquality <- read_csv("../../data/re_airquality/airquality.csv")}
```

## Data Overview

The data is a example data frame of "datasets" package. The database is a
secondary source. We must familiarize ourselves with its contents,
variables and units. For that we can use R to gather all these information.

### Missing values

With help() we see that the variables do not have the SI-units. With
sapply() we can check the class of the columns. With apply() we can see
that there are NA. We call a function to visualize the NA because we must
know where they are.

```{r Quality_control, error=FALSE, message=FALSE, warning=FALSE}
# Ask for details (units, source and references)
help("airquality")

# Check the class of each column
as.data.frame(sapply(airquality, class))

# Take a overview and main statistic parameters
summary(airquality)

# Check wheater there are any missing values (We try with only the first six rows)
head(apply(airquality,2, function(x) is.na(x)))

# Load the function
source("../../R/general/function.visualize.na.values.R")

# Function call to vis. the NA
visualize.na.values(airquality)
```

### Sources

The data frame shows different variables at different times. The
measurement series starts on 05.01.1973 and lasts until 09.30.1973. Each
variable has a value for each day (it could be "NA" as well). We
distinguish the variable according to the source.

Data for ozone were collected from the New York State Department of
Conservation.

-   Ozone [ppb] represent the average of the concentration between 1pm -
    3pm at Roosevelt Island, NYC.

Data for the meteorological data were collected from the US. National
Weather Service.

-   Wind [mph] is calculated as the average wind speed at 7am and 10am at
    La Guardia Airport, NYC

-   Temperature [degrees Fahrenheit] is the maximum daily temperature at la
    Guardia Airport, NYC

-   Solar. R [Langley] average in the frequency band 4000-7700 Angstroms
    (wavelength: 400-770 nano meter) from 8am to noon in Central Park NYC.

```{r editing_the_data_frame_in_SI_units, error=FALSE, message=FALSE, warning=FALSE}
# Load the function
source("../../R/re_airquality/function.unit.converter.R")

# Function call
airquality_mutated <- unit.converter(airquality)
```

## Statistics

Now that all units have been converted to SI units, the question about the
position, scattering, and distribution parameters arises. For the ozone
concentration is the median the right position parameter because it is
robust against outliers. This is important because a short-term elevated
ozone concentration will probably not endanger health. If we use the mean,
we could loss the overall picture due to a big one time outlier. If the
median has been used as a position parameter, it should be used the IQR as
a scattering parameter. This will allow us to estimate how high the data
dispersion is. But we are also interested in the maximal values. A
environment with a very high ozone concentration can be toxic (even a short
time exposure).

The table shows the results of the statistical parameters in a monthly
resolution. We round the results to integer numbers, because rational or
real numbers do not make sense in this context. Only a few data are
available for the month of Jun (frequency only 10 which means only a third
of the month is available). The median (May = 18 ppb, Jun = 23 ppb, Jul =
60 ppb, Aug = 52 ppb, Sept = 23 ppb) is always lower or equal to the
threshold value (60 ppm). Therefore, the air quality is acceptable. But
there are measurements in every month which exceed the threshold value
(maximum \> 60ppm). For Jul, Aug and Sep there are even more values (0.75
-quantile \> 60ppm). That means that there are isolated days, when the
ozone concentration is problematic. If we compare the standard deviation
and IQR we notice that they are not so different. If we also take into
account the span, we notice that the core of the data has a relatively
small scatter. In the 0.25 and 0.75 the scatter increases, especially for
the months Jul and Aug. Last but no least we can see that the median is
always smaller than the mean. We can conclude a positive skewness.

```{r statistical_parameters, error=FALSE, message=FALSE, warning=FALSE}
# Calculate the main statistic parameters.
knitr::kable(airquality_mutated|>
  group_by(Group)|>
  summarise(Frequncy = length(na.omit(Ozone)),
            Median = round(median(Ozone, na.rm = TRUE), digits = 0),
            IQR = round(IQR(Ozone, na.rm = TRUE), digits = 0),
            Q.25 = round(quantile(Ozone, 0.25, na.rm = TRUE), digits = 0),
            Q.75 = round(quantile(Ozone, 0.75, na.rm = TRUE), digits = 0),
            Max = round(max(Ozone, na.rm = TRUE), digits = 0),
            Min = round(min(Ozone, na.rm = TRUE), digits = 0),
            Mean = round(mean(Ozone, na.rm = TRUE),digits = 0),
            STD = round(sd(Ozone, na.rm = TRUE), digits = 0),
            Max = round(max(Ozone, na.rm = TRUE), digits = 0),
            Min = round(min(Ozone, na.rm = TRUE), digits = 0),
            Span = round(max(Ozone, na.rm = TRUE), digits = 0) - 
                   round(min(Ozone, na.rm = TRUE), digits = 0)))
```

### Box-Whisker-Plot

Probably it is difficult to imagine what the parameters mean for the data.
That is why we visualize them. We can see the same information (even more
if we add a jitter-plot) in a box-whisker-plot as in the table above. The
plot shows us that there are outliers (red triangle), but the boxes for
May, June and September are small which means the data scattering is small
(small IQR). The situation for July and August is different. The large box
indicates a large IQR for these months and the data scattering is
relatively high. Additional, the plot shows also the mean value. It is
higher than the median for each box. Therefore, the skewness is for all
months positive. Please note that there are only 9 measurements available
for June.

```{r boxplot_ozone_per_month, error=FALSE, message=FALSE, warning=FALSE}
# Load the function
source("../../R/re_airquality/function.boxplot.R")

# Function call
airquality.boxplot(airquality_mutated)
```

### Density-Plot

The density plot shows us that the data have a big span and a positive
skewness. But the core of the data (area with a high density) has an
acceptable span (which is basically the IQR). The graphs shows us also that
the density for the months May, June and September is bi-modal. That means,
that there are outliers (which we already see in the box-whisker-plot).
July and August are mono-modal but have a large span. This we could see in
the box-whisker-plot (large boxes) as well.

```{r density_ozone_per_month, error=FALSE, message=FALSE, warning=FALSE}
# Load the function
source("../../R/re_airquality/function.density.plot.R")

# Function call
density.airquality(airquality_mutated)
```

### Bar Plot and the distribution of the missing values

We detected many missing values. For a stringent analysis we need to know
which values are missing and where they are. Are there a systematic error
in the instrument (sensitivity)? Or simple forgot someone to write down the
measurement? For that we use a bar-plot which replace all NA values in -1.
Now we can easily see which values are missing. We are missing a lot of
values for June. To create new values through linear interpolation is not
an option because there is (1) to many NA and (2) the gap between to values
is to big. We would only falsify our analysis. But all other missing values
are not problematic and we can proceed straight forward.

```{r bar_plot_missing_ozone_per_month, error=FALSE, message=FALSE, warning=FALSE}
# Load the function
source("../../R/re_airquality/function.na.identifier.R")

# Function call
na.identifier(airquality_mutated )
```

### Correlation

Now we have a impression of our data. But we want to know, whether the
variables of the hypothesis are promisingly. For this purpose we create a
correlation matrix. Because we do not know if the ozone distribution is
linear or monotonic, we use the "Pearson" and the "Spearman" method. The
graphic shows us a high positive correlation between ozone concentration
vs. temperature (r = 0.7, $\rho = 0.77$ ) and a high negative correlation
between the ozone concentration and wind ( r = -0.6, $\rho = -0.59$ ).
There is also a positive correlation between the ozone concentration and
the solar radiation (r = 0.35, $\rho=0.35$) but not so strong. The
difference between the methods are not negligible. In this exercise, we
would use the "Spearman" method because it describes the correlation a
little bit better. Nevertheless, the chosen variables are promisingly.

```{r correlation_plot, error=FALSE, message=FALSE, warning=FALSE}

source("../../R/re_airquality/function.correlation.plot.R")

correlation.plot(airquality_mutated, c("pearson"))

correlation.plot(airquality_mutated, c("spearman"))
```

## Modelling

We want to be able to predict new values. For that we need a model. Here we
create a simple linear regression model. For that we calculate four
different models.

### Linear Regression Models

The first model is ozone vs. time. The model explains only about 2.5 % of
the variance (RSQ = 0.024). Both the coefficient and the intercept are not
significant (p-value \> 0.05) and we can not reject H0 (Is the p-value
insignificant, then it indicates there is insufficient evidence in the
sample to conclude that a non-zero correlation exists). We reject that
model.

The second model is ozone vs. temperature. The model explains about 48% of
the variance. That is not so bad but far from perfect. The p-value is \<
0.05 and therefore significant (even on a 99.9 % level). We can reject H0.
We accept the model temporarily.

The third model is ozone vs. wind. The RSQ is about 0.36. The model
explains about 36 % of the variance. The p-value is \< 0.05 and the
coefficient and the intercept are significant (even on a 99.9% level). We
reject H0 and accept the model temporarily.

The last model is ozone vs. solar radiation. This model is not as good as
the second and the third model. RSQ is only about 0.12 what explains only
12 % of the variance. The p-value is \< 0.05 but the intercept is only on a
99 % level significant (actually this is a good value but it is lower
compared to the second an third model). We reject H0 but the model is not
good. Because we have better option, we will not use this model.

```{r linear_regression_model_ozone_temperature, error=FALSE, message=FALSE, warning=FALSE}
# Load the function
source("../../R/re_airquality/function.regression.calculater.R")

# Function call ozone vs. Date
regression.calculater(airquality_mutated, 4, 1)

# Function call ozone vs. Temperature
regression.calculater(airquality_mutated, 4, 5)

# Function call ozone vs. Wind
regression.calculater(airquality_mutated, 4, 6)

# Function call ozone vs. Solar radiation
regression.calculater(airquality_mutated, 4, 7)
```

### Visualization of the models

For a better interpretation the models can be visualized.

```{r linear_regression_model_ozone_time, error=FALSE, message=FALSE, warning=FALSE}
# Load the function
source("../../R/re_airquality/function.regression.R")

# Function calls
airquality.regression(airquality_mutated, Day, c("Day"))

airquality.regression(airquality_mutated, Temp, c("Temp [K]"))

airquality.regression(airquality_mutated, Wind, c("Wind [m/s]"))

airquality.regression(airquality_mutated, Solar.R, c("Solar.R [W/m^2]"))
```

# Discussion

In this short analysis we can see that ozone probably depends on
temperature, wind and solar radiation. Although the RSQ and the p-value was
not bad for the second and third model, none of the models is convincing
enough to use as an explanatory model. We need better approaches. One
possibility would be multiple regression or machine learning approaches.
However, these are not used here because they will be introduced later.

Although the RSQ for the second and third model was not so bad, the
visualization shows a new problem. The RSQ in the model is valid for the
whole time-period. However, in the visualization (regression for each
month) you can see that the RSQ is highly fluctuating. It could be a good
idea, when we chose a new approach for each month. If we analyse more data
(maybe over 20-50 years) we could discover pattern. Perhaps we must use for
the month May another variable than for June. With only this data it is
impossible to say what we should do. If we look for patterns in this data,
any pattern found would probably be random.

But we can take a look on this specific data set and analyse the year 1973.
First we must discuss the source. It is problematic that not all variables
were measured at the same location. The ozone was measured on Roosvelt
Island. It is an island on the East River in NYC. But the temperature and
wind were measured at La Guardia Airfield, the solar irradiance in Central
Park. The measuring stations are up to 10 km apart and this is in the
middle of a city. There is no guarantee that the meteorological conditions
are the same. There is probably more wind at the airport than at Roosvelt
Island. It would be better to collect new, up-to-date data ourselves and
measure all variables at the same place. It would be even better if we
implemented an urban measurement network. This would probably improve the
model massively.

Second we can see, that the median is lower than the threshold for all
months. The temperature is a important factor for the ozone concentration.
The same is true for the wind. In the correlation plot showed us that there
is a correlation between temperature and wind (what makes sense in a
meteorological way). We also know, that the ozone concentration do not
correlate with time. This makes also sense. Although the ozone
concentration is higher during summer (summer smog), there is not relation
to the day of a month. The concentrations reacts on physical parameter like
radiation, wind and temperature. It makes absolutely sense, that the
temperature is higher Jun, Jul, Aug and Sept than May. The ozone
concentration tends to be higher, if the temperature is high. In a urban
space this effect will be amplified due to heat-islands. A city like NYC is
predestined for such an effect.

As a conclusion we can say that the air quality is less bad as expected.
But there are days, where the concentration is to high (true for all
months). The city of NY should take measures to prevent heat islands. This
could lower temperature, which inhibits ozone production. Further, the
source material should be limited. With less traffic, NOX could be lowered
what would inhibit the ozone production. But first we should collect more
data. We recommend to build up a city-wide monitoring network, which should
include more air quality parameters, such as particulate matter. This would
give us an overall picture and allow us to take evidence-based measures.
Unfortunately we could not answer our research question. We suppose it is
possible, if we collect more data. We are convinced (and our analysis seems
to support the thesis) that we are on the right path.

# Bibliography

Brönnimann Stefan (2013): Ozon in der Atmosphäre. Bern, Haupt Verlag. DOI:
10.480/GB2013.03
