---
title: "Report Exercise: re_airquality"
author: "Patrick Bigler"
date: "2023-04-10"
output: 
  html_document:
    toc: yes
    toc_float: yes
    number_sections: yes
editor_options: 
  markdown: 
    wrap: 75
---

Course: Applied Geo-Data Science at the University of Bern (Institute of
Geography)

Supervisor: Prof. Dr. Benjamin Stocker

Adviser: Dr. Koen Hufkens, Pepa Aran, Pascal Schneider

Further information: <https://geco-bern.github.io/agds/>

[Do you have questions about the workflow? Contact the Author:]{.underline}

Author: Bigler Patrick (patrick.bigler1\@students.unibe.ch)

Matriculation number: 20-100-178

Reference: Report Exercise 3 (Chapter 4)

# Introduction

## Objectives

Visualizations are a proper way to communicate the results. The variety of
diagrams, colors, and shapes is almost limitless. However, not all of them
are always equally suitable. It is essential to create the right graphic to
transport the message. With a realistic problem, the following goals should
be achieved:

-   Appropriate choice of visualizations for different data types and
    different aspects of the data

-   Grammar of graphics (ggplot2)

-   The proper use of colors in visualization

## Theory

Air quality depends on factors such as particulate matter, nitrogen dioxide
$(NO_2)$ or ozone $(O_3)$. Ozone is not a health hazard per se, but a high
concentration over a longer period of time can affect health (Brönnimann
2013, 23). But what are high concentrations? The limit value for hourly
mean ozone concentrations set in the "Swiss Air Pollution Control Act" is
$120\: \frac{\mu g}{m^3}$ (Brönnimann 2013, 23). However, this unit depends
on the air pressure, which decreases with altitude. The volume mixing ratio
with the unit parts per billion is more suitable, which is about 60 ppb
(Brönnimann 2013, 26). Here, we use this value.

Ozone is a molecule consisting of three oxygen atoms and has various
physical (absorption of UV and IR radiation) and chemical (high reactivity)
properties. Ozone emerges due to a two-phase chemical reaction (Brönnimann
2013, 75):

1.  $NO_2 + h\nu \left(\lambda < 420 nm \right) \rightleftharpoons NO + O$

2.  $NO + O_3 \rightleftharpoons NO_2 + O_2$

The reaction equations show that UV radiation, oxygen, and nitrogen dioxide
are needed for the formation of ozone. The first two are available in
almost unlimited quantities. Further, many chemical reactions are
temperature dependent. This is also true for ozone. The higher the
temperature, the more the balance of the reaction shifts towards ozone
production, and the concentration increases (Brönnimann 2013, 95).

In summary, the ozone concentration probably depends on temperature (shift
of the balance), solar radiation (mostly UV radiation), and the
concentration of nitrogen dioxide (which depends among other things on
wind). With these considerations, a data set on air quality can be studied
with a focus on ozone.

## Hypothesis

This exercise is supposed to answer the following research question:

*"Quantify the dependency of tropospheric ozone concentration on
temperature, solar radiation, and wind."*

# Methods

## Unit transformation

Using the SI unit system is strongly recommended because it simplifies the
reproducibility, interpretation, and discussion of the data. It follows the
formulas to transform the units into the SI-unit system.

-   The following formula was used for
    temperature:$$\frac{x \left[°Fahrenheit \right]- 32}{1.8} +273.15 \Longleftrightarrow x\left[Kelvin \right]$$

-   The following formula was used for wind:
    $$1 \left[\frac{mile}{hour} \right] \Longleftrightarrow 0.44704 \left[\frac{meter}{second}\right]$$

-   The following formula was used for radiant exposure
    $$1 \left[ Langley \right] \Longleftrightarrow 41840 \left[ \frac{Joule}{m^2}\right]$$

## R: Evaluation of the data

The open-source program R-Studio (Version 2022.12.0+353) was used for all
studies. The data handling in this program is done using the R language. We
utilize the R-package "Tidyverse" and other packages to process the data
effectively.

# Programming and data evaluation

## Packages

The following code chunk contains all packages we need. Important is the
package "conflicted". It enables us to chose if different functions have
the same call but do not make the same thing (a function in a certain
package can have the same name as a function from another package).

```{r Packages, error=FALSE, message=FALSE, warning=FALSE}
# Load all packages
source("../../R/general/packages.R")
```

## Read the file

First, we get the data and save it in our directory. But we must do this
only once. If the file exists, then we can read it directly. That is why we
implemented an if-else statement.

```{r Read_file, error=FALSE, message=FALSE, warning=FALSE}
# We want know, if a certain file already exists
name.of.file <- "../../data/re_airquality/airquality.csv"

# If do not exists such a file, create it!
if (!file.exists(name.of.file)){
  # Read the data
  airquality.raw <- datasets::airquality

  # Write a CSV file in the respiratory to save the data
  write_csv(airquality.raw, "../../data/re_airquality/airquality.csv")
  
  # Read the file
  airquality <- read_csv("../../data/re_airquality/airquality.csv")

  # If exists such a file, read it only!
  }else{airquality <- read_csv("../../data/re_airquality/airquality.csv")}
```

## Data Overview

The data is an example data frame from the "datasets" package. The database
is a secondary source. We must familiarize ourselves with its contents,
variables, and units. For that, we can use R to gather all the information.

### Missing values

With help() we see that the variables do not have SI units. With sapply()
we can check the class of the columns. With apply() we can see that there
are NA. We call a function to visualize the NA because we must know where
they are.

```{r Quality_control, error=FALSE, message=FALSE, warning=FALSE}
# Ask for details (units, source and references)
help("airquality")

# Check the class of each column
as.data.frame(sapply(airquality, class))

# Take a overview and main statistic parameters
summary(airquality)

# Check wheater there are any missing values (We try with only the first six rows)
head(apply(airquality,2, function(x) is.na(x)))

# Load the function
source("../../R/general/function.visualize.na.values.R")

# Function call to vis. the NA
visualize.na.values(airquality)
```

### Sources

The data frame shows different variables at different times. The
measurement series starts on May 1, 1973, and lasts until September 30,
1973. Each variable has a value for each day (it could be "NA" as well). We
distinguish the variable according to the source.

Data for ozone were collected from the New York State Department of
Conservation:

-   Ozone [ppb] represents the average concentration between 1 p.m. and 3
    p.m. at Roosevelt Island, NYC.

The meteorological data were collected from the US National Weather
Service:

-   Wind [mph] is calculated as the average wind speed at 7 a.m. and 10
    a.m. at La Guardia Airport, NYC.

-   Temperature [degrees Fahrenheit] is the maximum daily temperature at La
    Guardia Airport, NYC.

-   Solar. R [Langley] averages in the frequency band 4000--7700 Angstroms
    (wavelength: 400--770 nanometers) from 8 a.m. to noon in Central Park,
    NYC.

We want to use SI-units. That is why we make a function call and change
units into the SI system.

```{r editing_the_data_frame_in_SI_units, error=FALSE, message=FALSE, warning=FALSE}
# Load the function
source("../../R/re_airquality/function.unit.converter.R")

# Function call
airquality_mutated <- unit.converter(airquality)
```

## Statistics and its paramaters

Now that all units have been converted to SI units, the question of
position, dispersion, and distribution parameters arises. For the ozone
concentration, the median is a good position parameter because it is robust
against outliers. This is important because a short-term elevated ozone
concentration will probably not endanger health. If we use the mean as a
position parameter, we could lose the overall picture due to a big one-time
outlier. If the median has been used as a position parameter, the IQR
should be used as a dispersion parameter. This will allow us to estimate
how high the data scatter is. But we are also interested in the maximal
values. An environment with a very high ozone concentration can be toxic
(even for short-time exposure).

The table shows the results of the statistical parameters for a monthly
resolution. We round the results to integer numbers because rational or
real numbers do not make sense in this context. Only a few pieces of data
are available for the month of June (the frequency is only 10, which means
only a third of the month is available). The other months are better. The
median (May = 18 ppb, Jun = 23 ppb, Jul = 60 ppb, Aug = 52 ppb, Sept = 23
ppb) is always lower or equal to the threshold value (60 ppm). Therefore,
the air quality is not acceptable for almost every second day
(statistically speaking). There are measurements every month that exceed
the threshold value (maximum \> 60 ppm). For July, August, and September,
there are even more values (0.75 quantiles \> 60 ppm). That means that
there are isolated days when the ozone concentration is problematic.

If we compare the standard deviation and IQR, we notice that they are not
so different. If we also take the span into account, we notice that the
core of the data has a relatively small dispersion. In the range of 0.25
and 0.75, the scatter increases, especially for the months of July and
August. Last but not least, we can see that the median is always smaller
than the mean. We can conclude that there is a positive skewness.

```{r statistical_parameters, error=FALSE, message=FALSE, warning=FALSE}
# Calculate the main statistic parameters.
knitr::kable(airquality_mutated|>
  group_by(Group)|>
  summarise(Frequncy = length(na.omit(Ozone)),
            Median = round(median(Ozone, na.rm = TRUE), digits = 0),
            IQR = round(IQR(Ozone, na.rm = TRUE), digits = 0),
            Q.25 = round(quantile(Ozone, 0.25, na.rm = TRUE), digits = 0),
            Q.75 = round(quantile(Ozone, 0.75, na.rm = TRUE), digits = 0),
            Max = round(max(Ozone, na.rm = TRUE), digits = 0),
            Min = round(min(Ozone, na.rm = TRUE), digits = 0),
            Mean = round(mean(Ozone, na.rm = TRUE),digits = 0),
            STD = round(sd(Ozone, na.rm = TRUE), digits = 0),
            Max = round(max(Ozone, na.rm = TRUE), digits = 0),
            Min = round(min(Ozone, na.rm = TRUE), digits = 0),
            Span = round(max(Ozone, na.rm = TRUE), digits = 0) - 
                   round(min(Ozone, na.rm = TRUE), digits = 0)),
            caption = "Table 1: Overview of main statistical parameters (unit [ppb])",
            align = c("l", rep("c", 11)), format = "html")|>
  kableExtra::kable_classic()

```

### Box-Whisker Plot

But it is probably difficult to imagine what the parameters mean for the
data. That is why we visualize them. We can see the same information (even
more if we add a jitter-plot) in a box-whisker plot as in the table above.
The plot shows us that there are outliers (red triangle), but the boxes for
May, June, and September are small, which means the data dispersion is
small (small IQR). The situation for July and August is different. The
large box indicates a large IQR for these months, and the data dispersion
is relatively high. Additionally, the plot shows the mean value. It is
higher than the median for each box. Therefore, the skewness is positive
for all months. Please note that there are only nine measurements available
for June.

```{r boxplot_ozone_per_month, error=FALSE, message=FALSE, warning=FALSE}
# Load the function
source("../../R/re_airquality/function.boxplot.R")

# Function call
airquality.boxplot(airquality_mutated)
```

### Density-Plot

The density plot shows us that the data has a large span and a positive
skewness. But the core of the data (the area with highest density) has an
acceptable span (which is basically the IQR). The graphs also show us that
the density for the months of May, June, and September is bimodal which
means there are outliers (we already see that in the box-whisker plot).
July and August are monomodal but have a large span. This we could see in
the box-whisker plot (large boxes) as well.

```{r density_ozone_per_month, error=FALSE, message=FALSE, warning=FALSE}
# Load the function
source("../../R/re_airquality/function.density.plot.R")

# Function call
density.airquality(airquality_mutated)
```

### Bar Plot and the distribution of the missing values in a timeline

We detected many missing values. For a stringent analysis, we need to know
which values are missing and where they are. Is there a systematic error in
the instrument's sensitivity? Or simply forgot someone to write down the
measurement? For that, we use a bar plot that replaces all NA values with
-1. Now we can easily see which values are missing. We are missing a lot of
values for June. Creating new values through linear interpolation is not an
option because there are (1) too many NAs and (2) the gap between the
values is too big. We would only falsify our analysis. But all the other
missing values are not problematic, and we can proceed straight forward.

```{r bar_plot_missing_ozone_per_month, error=FALSE, message=FALSE, warning=FALSE}
# Load the function
source("../../R/re_airquality/function.na.identifier.R")

# Function call
na.identifier(airquality_mutated )
```

### Correlation

Now we have an impression of our data. But we want to know whether the
variables in the hypothesis are promising. We could create a correlation
plot for each variable-combination. But we would have a lot to do. For this
purpose, we created a correlation matrix. Because we do not know if the
ozone distribution is linear or monotonic, we use the "Pearson (r)" and
"Spearman ($\rho$)" method. The graphic shows us a high positive
correlation between ozone concentration and temperature (r =
0.7, $\rho = 0.77$) and a high negative correlation between ozone
concentration and wind ( r = -0.6, $\rho = -0.59$). There is also a
positive correlation between the ozone concentration and solar radiation (r
= 0.35, $\rho=0.35$) but it is not so strong. The difference between the
methods is in general not negligible. In this exercise, we will use the
"Pearson" method because we assume a linear relationship (we eventually
compute a linear regression model). Nevertheless, the chosen variables are
promising.

```{r correlation_plot, error=FALSE, message=FALSE, warning=FALSE}

source("../../R/re_airquality/function.correlation.plot.R")

correlation.plot(airquality_mutated, c("pearson"))

correlation.plot(airquality_mutated, c("spearman"))
```

We have seen that there is a negative correlation between temperature and
wind. From a meteorological point of view, this makes sense. That is why we
make a scatter plot where the wind has a temperature dependency. We see
that higher temperatures tend to produce lower winds. We also see that the
higher the wind, the higher the ozone concentration.

```{r wind_temp, error=FALSE, message=FALSE, warning=FALSE}
# Load the function
source("../../R/re_airquality/function.air.temp.R")

# Function call
temp.and.air(airquality_mutated)
```

## Modelling

We want to be able to predict new values (e.g. for interpolate the missing
values in June). For that, we need a model. Here, we create simple linear
regression models.

### Linear Regression Models

The first model is ozone vs. time. The model explains only about 2.5% of
the variance (RSQ = 0.024). Both the coefficient and the intercept are not
significant (p-value \> 0.05), and we cannot reject H0 (if the p-value is
insignificant, then it indicates there is insufficient evidence in the
sample to conclude that a non-zero correlation exists). We reject that
model.

The second model is ozone vs. temperature. The model explains about 48% of
the variance. That is not so bad, but it is far from perfect. The p-value
is \< 0.05 and therefore significant (even on a 99.9% level). We can reject
H0. We accept the model temporarily.

The third model is ozone vs. wind. The RSQ is about 0.36. The model
explains about 36% of the variance. The p-value is \<  0.05, and the
coefficient and the intercept are significant (even on a 99.9% level). We
reject H0 and accept the model temporarily.

The last model is ozone versus solar radiation. This model is not as good
as the second and third models. RSQ is only about 0.12, which explains only
12% of the variance. The p-value is \<  0.05, but the intercept is only at
a 99 % level of significance (actually, this is a good value, but it is,
lower compared to the second and third models). We reject H0, but the model
is not good. Because we have a better option, we will not use this model.

We set the code chunk to eval = FALSE because we do not want to call the
models. If you interest, you can set eval = TRUE and run the code again.

```{r linear_regression_model, eval=FALSE, error=FALSE, message=FALSE, warning=FALSE}
# Load the function
source("../../R/re_airquality/function.regression.calculater.R")

# Function call ozone vs. Date
regression.calculater(airquality_mutated, 4, 1)

# Function call ozone vs. Temperature
regression.calculater(airquality_mutated, 4, 5)

# Function call ozone vs. Wind
regression.calculater(airquality_mutated, 4, 6)

# Function call ozone vs. Solar radiation
regression.calculater(airquality_mutated, 4, 7)
```

### Visualization of the models

For a better interpretation, the models can be visualized.

```{r linear_regression_model_ozone_time, error=FALSE, message=FALSE, warning=FALSE}
# Load the function
source("../../R/re_airquality/function.regression.R")

# Function calls
airquality.regression(airquality_mutated, Day, c("Day"))

airquality.regression(airquality_mutated, Temp, c("Temp [K]"))

airquality.regression(airquality_mutated, Wind, c("Wind [m/s]"))

airquality.regression(airquality_mutated, Solar.R, c("Solar.R [W/m^2]"))
```

# Discussion

In this short analysis, we can see that ozone probably depends on
temperature, wind, and solar radiation. Although the RSQ and the p-value
were not bad for the second and third linear regression models, none of the
models is convincing enough to use as an explanatory model. We need better
approaches. One possibility would be multiple regression or machine
learning approaches. However, these are not used here because they will be
introduced later.

But we can take a look at this specific data set and analyze the year 1973.
First, we must discuss the source. It is problematic that not all variables
were measured at the same location. The ozone was measured on Roosevelt
Island. It is an island on the East River in NYC. But the temperature and
wind were measured at La Guardia Airfield, and the solar irradiance was
measured in Central Park. The measuring stations are up to 10 kilometers
apart, and this is in the middle of a city. There is no guarantee that the
meteorological conditions will be the same. It would be better to collect
new, up-to-date data ourselves and measure all variables in the same place.

Second, we can see that the median is lower or equal than the threshold for
all months. Temperature is an important factor in ozone concentration. The
same is true for wind. The ozone concentration tends to be higher if the
temperature is high. In an urban space, this effect will be amplified due
to heat islands. A city like NYC is predestined for such an effect.

In conclusion, we can say that the air quality is less bad than expected.
But it is still poor air quality. The median is often just not exceeded.
But there are still almost 50% of the days above the treshold value. There
are days when the concentration is too high (true for all months). The city
of New York should take measures to prevent heat islands. This could lower
the temperature, which inhibits ozone production. Further, the source
material should be limited. With less traffic, $NO_2$ could be lowered,
which would also inhibit ozone production. Unfortunately, we could not
answer our research question. We suppose it is possible if we collect more
data. We are convinced (and our analysis seems to support the thesis) that
we are on the right path.

(390 words)

# Bibliography

Brönnimann Stefan (2013): Ozon in der Atmosphäre. Bern, Haupt Verlag. DOI:
10.480/GB2013.03
