---
title: "Exercise_3"
author: "Patrick Bigler"
date: "2023-04-24"
output:
  pdf_document: default
  html_document: default
editor_options: 
  markdown: 
    wrap: 75
---

# Report Exercise: "Stepwise regression"

Course: Applied Geo-data Science at University of Bern (Institute of
Geography)

Supervisor: Prof. Dr. Benjamin Stocker

Adviser: Dr. Koen Hufkens, Pepa Aran, Pascal Schneider

[You have questions to the workflow? Contact the Author:]{.underline}

Author: Bigler Patrick (patrick.bigler1\@students.unibe.ch)

Matriculation number: 20-100-178

Reference: Report Exercise 3 (Chapter 8)

## Introduction

This exercise demonstrates an algorithm to select the best possible linear
regression model. In R, there are functions for this that allow a fast and
efficient solution. Here, however, the whole algorithm is to be implemented
as an exercise. Roughly speaking, the algorithm should do the following:

1.  Set the number of predictors to be considered to p equal 1.

2.  Fit all regression models with p predictors and compute their R\^2.

3.  Select the model with p predictors that achieves the highest R\^2 (best
    fitting model) and compute its AIC.

4.  Increment to p + 1. Fit all regression models with p + 1 predictors and
    compute their R\^2. Select the best fitting model and compute its AIC.

5.  If the AIC of the model with p + 1 predictors is poorer than the AIC of
    the model with p predictors, retain the model with p predictors and
    quit.

6.  You have found the (presumably) optimal model. Otherwise, continue with
    with step 4.

For the first task we apply only the first 3 steps. If we do that we may be
able to chose the best bivariate model. For the second task we iterate the
steps 4 and 5. If we do that we might also be able to chose the best
multivariate model. Please note that the functions are outsourced to a
separated R-script.

After we implemented the algorithm you will see, that there are a lot to
discuss. Although the algorithm is clear, the interpretation and the
application may not.

## Packages

The following code chunk contains all packages we need. Important is the
package "conflicted". It enables us to chose if different functions have
the same call but do not make the same thing (a function in a certain
package can have the same name as a function from another package).

```{r Packages, error=FALSE, message=FALSE, warning=FALSE}
source("../../R/general/packages.R")
```

First, we get the data and save it in our repository. Because we can not be
sure if there made any changes in the data, we "freeze" our data. That
means we get the data, save it and read it from our directory.

```{r Access_to_the_data eval=FALSE}
# Access to the data
url <- "https://raw.githubusercontent.com/geco-bern/agds/main/data/df_for_stepwise_regression.csv"

# Read in the data directly from URL
database.S1 <- read.table(url, header = TRUE, sep = ",")

# Write a CSV file in the respiratory
write_csv(database.S1,"../../data/re_stepwise/FLX_CH_stepwise.csv")
```

To make a good decision about the predictors we need as much information as
possible. For that we read our file as a data frame. After that, we want a
overview about the data ( what variables contains teh data set? What are
the main statistical parameters?)

```{r Read_the_data, error=FALSE, message=FALSE, warning=FALSE}
database <- read_csv("../../data/re_stepwise/FLX_CH_stepwise.csv")

# Take a overview and main statistic parameters
summary(database)
```

Now we know the magnitude and we gain some information about the
distribution. But it is very difficult to decide if a linear regression is
an appropriate approach. To solve this problem, we calculate a overview
about all simple linear regression models and plot it. Now we can see that
for most of the variables a linear regression is a good idea.

```{r Overview_about_the_data,error=FALSE, message=FALSE, warning=FALSE}
# Overview about the data, the regression

database|>
  pivot_longer(cols = c(database|>
  select(-c("siteid", "TIMESTAMP", "GPP_NT_VUT_REF")) |>
  colnames()),
               names_to = "Variable", 
               values_to = "Value")|>
  group_by(`Variable`)|>
  ggplot(aes(x = `Value`, y = `GPP_NT_VUT_REF`, na.rm = TRUE)) +
  geom_point(alpha = 0.5) +
  geom_smooth(formula = y~x, method = "lm", color = "red", se = FALSE) +
  facet_wrap(~Variable )
```

But which is probably the best one? With which variable should we start?
For that a correlation plot is a good thing. It gives us a good overview
about the correlation. In the plot before, we have also seen that the
relation of the variables is mostly linear. So we use the "pearson" method
to calculate the correlation.

```{r Visualization_of_the_data,error=FALSE, message=FALSE, warning=FALSE}
# create a correlation matrix. Use "use" to ignore NAs
cor.mat <- cor(database[, 3:17], use = "complete.obs", method = "pearson")

# plot a correlation plot. Now we have a good overview
ggcorrplot(cor.mat, hc.order = TRUE, type  = "lower",
   lab = FALSE)
```

Now we have gain many information about the variables. But it could be very
difficult to decide which variable are the best for a goof lm fit. For that
we use a function which will make this decision for us. If we want only one
predictor to describe the target we use a bivariate linear regression
model.

## Bivariate linear regression

```{r Calculate_best_Bivariante_Model, error=FALSE, message=FALSE, warning=FALSE}
# Access the outsourced function for the bivariate model
source("../../R/re_stepwise/function.bivariate.model.R")

# Function call
tibble.bi.model <- model.fitter(database, 16)
```

### Visualization of the bivariate model

There are meaningful scatter-plot for bivariate lm -models. It gives us a
good first impression. But we want to know more about the coefficients, the
intercept, the distribution and outliers. For that we use summary() and we
plot the model.

```{r Vis_bivariate_regression_model, error=FALSE, message=FALSE, warning=FALSE}
# Calculate the probebly best multivariate model
best.bi.lm <- lm(GPP_NT_VUT_REF~PPFD_IN, data = database)

# Load the function
source("../../R/re_stepwise/function.vis.bi.model.R")

# Plot the bivariate model
vis.bi.model(database)

# Overview about all important model parameter
# Model Call. We are also interested for the p-values
summary(best.bi.lm)

# Plot the model 
plot(best.bi.lm)
```

```{r Vis_RSQ_development, error=FALSE, message=FALSE, warning=FALSE}

# Load the function
source("../../R/re_stepwise/function.vis.and.dev.bi.R")

# Function call
Vis_of_the_develop.bi.model(tibble.bi.model)
```

## Multivariate linear regression

```{r best_multivariate_model, error=FALSE, message=FALSE, warning=FALSE}
# Access the outsourced function of the multivariate model
source("../../R/re_stepwise/function.multivariate.model.R")

# Function call with all column but target (column number 16)
multivariate.model(database, c(16))

# Function call without the column numbers 1, 2 and 16
tibble.multi.model <- multivariate.model(database, c(1,2,16))
```

### Visualization of the multivariate model

There are no meaningful scatter-plot for multidimensional models. But we
want to know more about the coefficients, the intercept, the distribution
and outliers.

```{r Visn_multivariate_regression_model, error=FALSE, message=FALSE, warning=FALSE}
# Calculate the probably best multivariate model
best.multi.lm <- lm(GPP_NT_VUT_REF~PPFD_IN + LW_IN_F + VPD_F + TA_F_MDS + 
                      SW_IN_F + P_F + WS_F + CO2_F_MDS + PA_F, data = database)

# Overview about all important model parameter
# Model Call
summary(best.multi.lm)

# Plot the model 
plot(best.multi.lm)
```

```{r Vis_RSQ_AIC_development, error=FALSE, message=FALSE, warning=FALSE}
# Load the function
source("../../R/re_stepwise/function.vis.and.dev.multi.R")

# Function call
vis.and.dev.multi.model(tibble.multi.model)
```

## Discussion

It seems that the step forward algorithm is successfully implemented.
However, it is important to choose the predictors very carefully because
the final model could varies. The target, of course, must be removed before
we can use the algorithm because if we use the target as a predictor, R\^2
will be always 1. Additional, the predictor would be a linear combination
of the target and therefore disrupt the results.

All other column must be chosen very carefully as well. For that, we have
to understand for what the column stands. After that we must decide whether
the variable is a meaningful predictor. For example, could be time very
important. Maybe the experiment-setting were outdoor and the climate
factors were important. It could also be possible, that our experiment
setting were indoor and there were always labor-conditions. That means,
that the date and time is irrelevant. It is not always clear what is a
"good choice".

For task 1 I could have wrote a very similar function as for task two. But
I wrote a function which is a little bit complicated as necessary. I did
this to demonstrate how the algorithm works. The algorithm calculate a
linear regression model for each predictor. For the model with the highest
R\^2 w. You can see on the table that the the model with the highest model
is not necessary the model with the lowest AIC.

For task two I wrote also a function. It is a step forward algorithm.
First, the function calculate a bivariate linear regression model and chose
the one which has the highest R\^2. After that, a multivariate regression
model is calculated and again the one with the highest R\^2 is chosen. For
each round, we compare the AIC. If the AIC higher than the model before, we
stop the calculation and we have found our probebly best fit model. But
here we have the same problem as describe above. Our calculation depends on
our input. We therefore need to consider which variables to include in the
calculation.

To demonstrate that, I made different function calls. You can easely see
that the results are different. That means we must (1) chose wise (maybe
with educational guess) (2) try different calls to be able to estimate how
big the difference is and (3) document how and why what was decided.

### Discussion of plot(model) call

Inspired by: <https://www.statology.org> and the book: "Statistik" form L.
Fahrmeir

#### For the bivariate model

Residuals vs. fitted plot We can see the right line follows almost the
dashed line. It is near to zero and approximately a horizontal line. That
means the linearity is not violated and the assumption that we use a linear
regression model were ok.

Q-Q-Plot The Q-Q-Plot shows us that the residuals are almost following a
normal distribution. The error is probably mainly statistical and not
systematic. The plot shows also that the choice of a linear model was the
right one because it follows almost the line with a 45° angle (It would
then the values has a perfect normal distribution).

Scale-Location Plot The red line should be roughly horizontal across the
plot. If it so, then the assumption of homoscedasticity is probably right.
The scattering should be random and there should be no pattern at all. Both
are approximately true. The spread of the residuals is roughly equal at all
fitted values.

Leverage-Plot There are none values behind the cook's distance. That means
that there are no significant outliers which distort our analysis.

#### For the multivariate model

The plots are approximately the same as for the bivariate model and need no
more discussion. Only the scale-location plot needs a few words: The red
line is not horizontal. That means there could be a problem with the
homoscedasticity. But is is not so dramatically and need an
intervention.But we should keep this in mind.

### Development of R\^2 and AIC

We can see that R\^2 increases fast for the first few steps. But then it
slows down and suddenly, there is no further improvement and we can stop
our model. For the AIC we can see a similar behavior. But AIC decreases
instead of increasing. First, the decreasing is fast and slows down. In the
last step the AIC increased a bit. That is the time we stop our
calculation.

AIC is a kind of quality measure for the models. It is based on
log-likelihood and a correction term that takes into account the number of
variables. At the beginning, AIC becomes smaller. When the influence of the
correction term becomes too large, the AIC does not get smaller anymore and
we have a basis for decision making for the model choice.

Actually, we should discuss the adjusted RR. But this is based on RR and it
does not matter which one we use. Our model choice is ultimately based on
the AIC. After the model selection, we then have to select the adjusted RR.

The RR is based on the variance decomposition theorem, which states that
the variance of the dependent variable can be written as the sum of a part
of the variance explained by the regression model and the variance of the
residuals (unexplained variance). By definition, it increases with
additional variables, but the growth flattens out.

It is also important to note that RR does not simply add up, but must be
constantly recalculated. This means that for a multivariate model, we
cannot simply use the highest RR of the individual bivariate models or the
variables with the highest correlation coefficients. This is visible in the
figures "visualization of the development of RR and AIC (multi and bi). We
can see that the variable with the second best RR is SW_IN_F. But if we
improve our bivariate model to a multivariate model we do not use SW_IN_F
but LW_IN_F. But LW_IN_F only ranked eighth for a bivariate model. That why
we have to use the AIC for a goof model choice and that is also why we
recalculate the RR in our step forward algorithm in each round.
