---
title: "Report Exercise: re_stepwise"
author: "Patrick Bigler"
date: "2023-04-24"
output:
  html_document:
    toc: yes
    toc_float: yes
    number_sections: yes
editor_options: 
  markdown: 
    wrap: 75
---

Course: Applied Geo-data Science at University of Bern (Institute of
Geography)

Supervisor: Prof. Dr. Benjamin Stocker

Adviser: Dr. Koen Hufkens, Pepa Aran, Pascal Schneider

Further information: <https://geco-bern.github.io/agds/>

[You have questions to the workflow? Contact the Author:]{.underline}

Author: Bigler Patrick (patrick.bigler1\@students.unibe.ch)

Matriculation number: 20-100-178

Reference: Report Exercise 3 (Chapter 8)

# Introduction

## Objectives

Modelling is a important aspect of data sciences. But there are different
types of models with different complexity. With a realistic problem, the
following goals should be trained:

-   Understand the basics of regression and classification models.

-   Fit linear and logistic regression models in R.

-   Choose and calculate relevant model performance metrics.

-   Evaluate and compare regression models.

-   Detect data outliers.

-   Select best predictive variables.

## Theory

In environmental and climate science, dozens of variables have been
measured. With a model we want to try to predict a target. For that we use
a linear model. Unfortunately, we do not know how many variables and which
variables we should use. We only know that we try to find a formula like
this:

$$
\hat{y}=a_0 +\sum_{n=1}^{N}\hat{a}_n*x_n
$$

# Methods

To build up the best linear regression model we could, of course, simply
try all combinations by permutation. But that would be a waste of
resources. However, we cannot avoid a try and error approach. But we would
like to reduce the number of combinations. We implement a algorithm which
works as followed:

1.  Let $p\in[1,N]$ and a predictor. Let $N\in[1,max(variable)]$. N could
    be all variables we have or only a subset of it. Our algorithm will
    decide this.

2.  The algorithm takes p and create a linear model. The p with the highest
    RSQ will be used as a predictor.

3.  The algorithm takes p+1 and create a linear model. The p with the
    highest RSQ will be used as a predictor. Select the model with the
    highest RSQ and calculate the AIC.

4.  Iterate step 3 until the AIC of the model do not decreasing anymore.

5.  We have found the (presumably) optimal model.

For the first task we apply only the first 2 steps because we want only a
bivariate model. For the second task we iterate the step 3. If we do that
we might also be able to chose the best multivariate model.

## R: Evaluation of the data

All analysis are done with the open source software R-Studio (Version
2022.12.0+353). This software use the language R to handle the data. For an
efficient processing of the data we use the R-package "Tidyverse (and
others).

# Programming and data evaluation

## Packages

The following code chunk contains all packages we need. Important is the
package "conflicted". It enables us to chose if different functions have
the same call but do not make the same thing (a function in a certain
package can have the same name as a function from another package).

```{r Packages, error=FALSE, message=FALSE, warning=FALSE}
source("../../R/general/packages.R")
```

## Read the file

First, we get the data and save it in our repository. But we must do this
only once. If the file exists, then we can read it directly. That is why we
implemented a if-else statement.

```{r Read_file, error=FALSE, message=FALSE, warning=FALSE}
# We want know, if a certain file already exists
name.of.file <- "../../data/re_stepwise/FLX_CH_stepwise.csv"

# If do not exists such a file, create it!
if (!file.exists(name.of.file)){
  # Access to the data
  url <- "https://raw.githubusercontent.com/geco-bern/agds/main/data/df_for_stepwise_regression
  .csv"
  
  # Read in the data directly from URL
  database.S1 <- read.table(url, header = TRUE, sep = ",")

  # Write a CSV file in the respiratory
  write_csv(database.S1,"../../data/re_stepwise/FLX_CH_stepwise.csv")
  
  # Read the file
  database <- read_csv("../../data/re_stepwise/FLX_CH_stepwise.csv")

  # If exists such a file, read it only!
  # Read the file
  }else{database <- read_csv("../../data/re_stepwise/FLX_CH_stepwise.csv")}
```

## Data Overview

### Missing values

To make a good decision about the predictors we need as much information as
possible. For that we read our file as a data frame. After that, we want a
overview about the data ( what variables contains teh data set? What are
the main statistical parameters?)

```{r Read_the_data, error=FALSE, message=FALSE, warning=FALSE}
# Check wheater there are any missing values (We try with only the first six rows)
head(apply(database,2, function(x) is.na(x)))

# Load the function
source("../../R/general/function.visualize.na.values.R")

# Function call to vis. the NA
visualize.na.values.without.groups(database)

# Take a overview and main statistic parameters
summary(database)
```

## Bivariate linear regression

We chose the best variable for our predictor. For that, we try all
variables and chose the one with the highest RSQ (see introduction)

```{r Calculate_best_Bivariante_Model, error=FALSE, message=FALSE, warning=FALSE}
# Access the outsourced function for the bivariate model
source("../../R/re_stepwise/function.bivariate.model.R")

# Function call
tibble.bi.model <- model.fitter(database, 16)
knitr::kable(tibble.bi.model)
```

### Visualization of the bivariate model

There are meaningful scatter-plot for bivariate lm -models and it gives us
a first impression. But we want to know more about the coefficients, the
intercept, the distribution and outliers. For that we use summary() and we
plot the model (see discussion).

```{r Vis_bivariate_regression_model, error=FALSE, message=FALSE, warning=FALSE}
# Calculate the probebly best multivariate model
best.bi.lm <- lm(GPP_NT_VUT_REF~PPFD_IN, data = database)

# Load the function
source("../../R/re_stepwise/function.vis.bi.model.R")

# Plot the bivariate model
vis.bi.model(database)

# Overview about all important model parameter
# Model Call. We are also interested for the p-values
summary(best.bi.lm)

# Plot the model 
plot(best.bi.lm)
```

### Development of RSQ

What is happen with RSQ? Here we visualize the RSQ for each Variable. We
can easely see that PPFD_IN has the highest RSQ and that is why we chose it
as our predictor. But what if we want a multivariate regression? One might
think that we should simply choose the variable with the second highest
RSQ. But we will see that this is generally not the case.

```{r Vis_RSQ_development, error=FALSE, message=FALSE, warning=FALSE}

# Load the function
source("../../R/re_stepwise/function.vis.and.dev.bi.R")

# Function call
Vis_of_the_develop.bi.model(tibble.bi.model)
```

## Multivariate linear regression

Here we want to answer the question which predictors we should choose to
get the probably best linear regression model. For that we have wrote a
function. It works as described in the introduction. We make two function
calls. For the first we use all variables but the target. This is, however,
not necessarily the best way. We could say that the time factor and the
location is irrelevant and we drop out these variables. We can see that the
difference is huge.

```{r best_multivariate_model, error=FALSE, message=FALSE, warning=FALSE}
# Access the outsourced function of the multivariate model
source("../../R/re_stepwise/function.multivariate.model.R")

# Function call with all column but target (column number 16)
multi.model.1 <- multivariate.model(database, c(16))

# Function call without the variables siteid, TIMESTAMP and GPP_NT_VUT_REF
multi.model.2 <- multivariate.model(database, c(1,2,16))

# Overview about the models
knitr::kable(multi.model.1)
knitr::kable(multi.model.2)
```

### Development of RSQ and AIC

We visualize the development of the model. We can see that the model with
more variables has a higher RSQ. That is not a surprise because RSQ will
always increase if we add a variable. Important is the AIC. We can see that
the AIC is lesser for more variables (21762.2 vs. 24469.3). The RSQ is
higher (0.6579 vs. 0.5983). It follows that model 1 is better than model 2.
We see that also in the coefficients.

```{r Vis_RSQ_AIC_development, error=FALSE, message=FALSE, warning=FALSE}
# Load the function
source("../../R/re_stepwise/function.vis.and.dev.multi.R")

# Function call
vis.and.dev.multi.model(multi.model.1)
vis.and.dev.multi.model(multi.model.2, c("[without siteid and TIMESTAMP]"))
```

### Visualization of the multivariate model

There are no meaningful scatter-plot for multidimensional models. But we
want to know more about the coefficients, the intercept, the distribution
and outliers. We make some calls by using summary(). We get the same values
for the RSQ. This is proof that our algorithm works. We see that not all
coefficients are significant.

```{r Visn_multivariate_regression_model, error=FALSE, message=FALSE, warning=FALSE}
# Calculate the probably best multivariate model
best.multi.lm <- lm(GPP_NT_VUT_REF~PPFD_IN + LW_IN_F + VPD_F + TA_F_MDS + 
                      SW_IN_F + P_F + WS_F + CO2_F_MDS + PA_F + siteid + TIMESTAMP, 
                    data = database)

second.multi.lm <- lm(GPP_NT_VUT_REF~PPFD_IN + LW_IN_F + VPD_F + TA_F_MDS + 
                      SW_IN_F + P_F + WS_F + CO2_F_MDS + PA_F, 
                    data = database)

# Overview about all important model parameter
# Model Call
summary(best.multi.lm)

summary(second.multi.lm)

# Plot the model 
plot(best.multi.lm)
```

# Discussion

## General

It seems that the step forward algorithm is successfully implemented.
However, it is important to choose the predictors very carefully because
the final model could varies. The target, of course, must be removed before
we can use the algorithm because if we use the target as a predictor, RSQ
will be always 1. Additional, we must be aware of linear combinations. We
must always know for what each variable stands for. Linear combination
could disrupt our results. After that, we must decide whether the variable
is a meaningful predictor. For example: a time-factor could be very
important. Maybe the experiment-setting were outdoor and the climate
factors were important. It could also be possible, that our experiment
setting were indoor and there were always labor-conditions. That means,
that the date and time is irrelevant. It is not always clear what is a
"good choice".

For task 1 we could have wrote a very similar function as for task two. But
we wrote a function which is a little bit complicated as necessary. We did
this to demonstrate how the algorithm works. The algorithm calculate a
linear regression model for each predictor. With the table we are able to
see that the model with the highest RSQ is not necessary the model with the
lowest AIC.

For task two we wrote also a function. It is a step forward algorithm.
First, the function calculate a bivariate linear regression model and chose
the one which has the highest RSQ. After that, a multivariate regression
model is calculated and again the one with the highest RSQ is chosen. For
each round, we compare the AIC. If the AIC higher than the model before, we
stop the calculation and we have found our probably best fit model. But
here we have the same problem as describe above. Our calculation depends on
our input. Therefore, we need to consider which variables we include in the
calculation.

For demonstration, we made some function calls. We can easily see that the
results are different. That means we must (1) chose wise (maybe with
educational guess) (2) try different calls to be able to estimate how big
the difference is and (3) document how and why we have what decided.

## Discussion of plot(model) call

Inspired by: <https://www.statology.org> and the book: "Statistik" form L.
Fahrmeir

### For the bivariate model

Residuals vs. fitted plot

We can see the right line follows almost the dashed line. It is near to
zero and approximately a horizontal line. That means the linearity is not
violated and the assumption that we use a linear regression model were ok.

Q-Q-Plot

The Q-Q-Plot shows us that the residuals are almost following a normal
distribution. The error is probably mainly statistical and not systematic.
The plot shows also that the choice of a linear model was the right one
because it follows almost the line with a 45° angle (It would then the
values has a perfect normal distribution).

Scale-Location Plot

The red line should be roughly horizontal across the plot. If it so, then
the assumption of homoscedasticity is probably right. The scattering should
be random and there should be no pattern at all. Both are approximately
true. The spread of the residuals is roughly equal at all fitted values.

Leverage-Plot

There are none values behind the cook's distance. That means that there are
no significant outliers which distort our analysis.

### For the multivariate model

The plots are approximately the same as for the bivariate model and need no
more discussion. Only the scale-location plot needs a few words: The red
line is not horizontal. That means there could be a problem with the
homoscedasticity. But is is not so dramatically and do not need an
intervention.But we should keep this in mind.

## Development of RSQ and AIC

We can see that RSQ increases fast for the first few steps. But then it
slows down and suddenly, there is no further improvement and we can stop
our model. For the AIC we can see a similar behavior. But AIC decreases
instead of increasing. First, the decreasing is fast and slows down. In the
last step the AIC increased a bit. That is the time we stop our
calculation.

The RSQ is based on the variance decomposition theorem, which states that
the variance of the dependent variable can be written as the sum of a part
of the variance explained by the regression model and the variance of the
residuals (unexplained variance). By definition, it increases with
additional variables, but the growth flattens out. Actually, we should
discuss the adjusted RSQ insted of RSQ. RSQ is always a bit higher than the
adjusted RSQ. But it does does not matter which one we use in the
algorithm. The algorithm stops, if the AIC increase. If we chose a final
model, then we have to use the adjusted RSQ. AIC is a kind of quality
measure for the models. It is based on log-likelihood and a correction term
that takes into account the number of variables. At the beginning, AIC
becomes smaller. When the influence of the correction term becomes too
large, the AIC does not get smaller anymore and we have a basis for
decision making for the model choice.

It is also important to note that RSQ does not simply add up, but must be
recalculated constantly. This means that for a multivariate model, we
cannot simply use the highest RR of the individual bivariate models or the
variables with the highest correlation coefficients. This is visible in the
figures "visualization of the development of RR and AIC (bivariate). That
why we have to use the AIC for a good model choice and that is also why we
recalculate the RR in our step forward algorithm in each round.
