---
title: ""
output:
editor_options:
  markdown:
    wrap: 75
---

# Report Exercise: "Data wrangling with tidyverse"

Course: Applied Geo-data Science at University of Bern (Institute of
Geography)

Supervisor: Prof. Dr. Benjamin Stocker

Adviser: Dr. Koen Hufkens, Pepa Aran, Pascal Schneider

[You have questions to the workflow? Contact the author:]{.underline}

Author: Bigler Patrick (patrick.bigler1\@students.unibe.ch)

Matriculation number: 20-100-178

Reference: Report Exercise 1 (Chapter 3)

## Introduction

The topic of this exercise is the "Logarithmic Respond Ratio" (LRR). The
LRR is the natural logarithm ($log()$)) of the response ratio. The ratio
describes the elevated concentration of a quantity (e.g. CO2 or nitrogen)
to the ambient concentration and is defined as followed:

$$LRR = \log\left(\frac{x_{eluvated}}{x_{ambient}}\right)$$

In this exercise, we work with a data set which contains data of soil
organic carbon measurements in experiments where ecosystems are exposed to
ambient (low) and elevated (high) CO2 concentrations. The mean soil organic
carbon of multiple samples is recorded within each experiment for different
sample dates. Information is provided for the time in years since the start
of the experiment.

### Original source of the file

The data originate from a paper of Groenigen et al (2014). In this
exercise, we take a close look to the $CO_2$ - concentrations
$(\: [\frac{g}{m^{-2}}])$ in the experiments and calculate the LRR. Because
we want a comprehensive analysis of the data, we calculate the LRR in
different ways. For the scientific reproducibility, you will find the
original source is the following article (or in the respiratory):

*Groenigen, Kees Jan van, Xuan Qi, Craig W. Osenberg, Yiqi Luo, and Bruce
A. Hungate. "Faster Decomposition Under Increased Atmospheric CO2 Limits
Soil Carbon Storage." Science 344, no. 6183 (May 2, 2014):
508--9. <https://doi.org/10.1126/science.1249534>.*

## Method

This section is used to briefly introduce the methods. It will be shown
what software was used and which edits were made by hand.

### Editing the file manually

Unfortunately, the original file is not well structured. It has several
headings and empty spaces which complicates to read the file and therefore,
we must the file edit manually. The software EXCEL (MS Office 360 with the
license from the University of Bern, status: 2023-05-07) is used for this
task. The following points were changed manually:

1.  Filled the gaps between two Experiments with the name of the
    experiment.

2.  Removed the header and all comments

### Method and evaluation of the data

After the manual editing, the xlsx file can be read with the software
R-Studio (Version 2022.12.0+353). This software use the language R to
handle the data. In this part of the report it is shown how the data was
read and processed with the software. For an efficient processing of the
data we use the R-package "Tidyverse (and others).

## Programming and data evaluation

### Read the csv file and get a overview of the data and data controlling

The following code chunk contains all packages we need. Important is the
package "conflicted". It enables us to chose if different functions have
the same call, but do not make the same thing (a function in a certain
package can have the same name as a function from another package). In this
case, we will set preferences.

```{r Packages, error=FALSE, message=FALSE, warning=FALSE}
# Load the packages
source("../../R/general/packages.R")
```

The following code chunk shows how we can read the file. First, the
xlsx-file has several sheets. We extract the sheet of interest and save it
on a variable. But we want a csv-file and so we create a new folder and
save the new file into it. We read the csv-file again for further editing.
But we must do this only once. If the file exists, then we can read it
directly. That is why we implemented a if-else statement.

```{r error=FALSE, message=FALSE, warning=FALSE}
# We want know, if a certain file already exists
name.of.file <- "../../data/re_tidy.data/re_tidy.new.data.csv"

# If do not exists such a file, create it!
if (!file.exists(name.of.file)){
  # Read a certain sheet of the xlsx file
  raw <- read_xlsx("../../data-raw/re_tidy.raw.data/data.groenigen.xlsx", 
         "Database S1",range = NULL, col_names = TRUE, col_types = NULL, na = "")

  # convert the xlsx to a csv file
  write_csv(raw, "../../data/re_tidy.new.data.csv")
  
  # Read the file
  raw.data <- read_csv("../../data/re_tidy.data/re_tidy.new.data.csv", 
              show_col_types = FALSE)

  # If exists such a file, read it only!
  }else{raw.data <- read_csv("../../data/re_tidy.data/re_tidy.new.data.csv", 
                    show_col_types = FALSE)}
```

Information about the data is key. We make some calls for quality-control.
We change the names of the columns we need. Now, we can refer to them
better. Furthermore, we want to know which class the columns are and
weather there are missing values. We can see there are many NA (23.6%). The
horizontal lines implies that there are empty rows. But this is not a
problem at all. R will ignore them. We are good to go straight forward.

```{r Quality_control, error=FALSE, message=FALSE, warning=FALSE}
# Shows the first 6 rows (column names!)
head(raw.data)

# Check the class of each column
as.data.frame(sapply(raw.data, class))

# Return the important statistical parameters
summary(raw.data)

# change the names of the two columns that are needed later on
raw.data <- raw.data|>
  rename("mean ambient CO2" = `ambient CO2...6`,
         "mean increased CO2" = `increased CO2...7`)

# Load the function
source("../../R/general/function.visualize.na.values.R")
# Function call
visualize.na.values.without.groups(raw.data)
```

### Theory about how to calculate the LRR for each experiment

For the first task we calculate the LRR for each experiment. But we can do
this in two different ways:

Method 1: Firstly, we group the database by 'Experiment'. Secondly, we
calculate the mean for the columns "mean ambient CO2" and "mean increased
CO2" (for each experiment). Thirdly, we calculate the LRR with the values.

Method 2: Firstly, we calculate the LRR for each row. Secondly, we group
the database by "Experiment". Thirdly, we take the mean of all LRR of each
experiment.

Although the results of both methods are the LRR for each experiment, they
will be in general different. Therefore, it is important which method we
chose. In this case, it is more meaningful to use method 2 because every
experiment has a sample date. In the experimental setup are the CO2
concentrations fixed. It makes more sense that we take this into account
and calculate the LRR for a specific sample date. If we use method 1, we
would calculate LRR with a CO2 concentration which never existed in the
experimental setup. With method 2 we use concentrations who has been
measured in an experiment and it makes therefore more sense. Here we will
calculate the LRR with both methods because we want to demonstrate, that
there is a difference.

### Theory about how to calculate the LRR for each phase

In this second task the experiments are divided into three phases,
accounting the duration of the experiment.

1)  Phase "early" $\Longrightarrow$ lesser than 3 years

2)  Phase "med" $\Longrightarrow$ between 3 and 6 years

3)  Phase "late" $\Longrightarrow$ greater than 6 years

For an efficient way to processing the data we use "Tidyverse" again. Here,
we are confronted with the same problem as in the first task because there
are again the two methods to calculate the LRR. We chose method 2 again
(same considerations and the same argumentation as for task one). We
calculate the LRR with both methods because we want to demonstrate that
there is a difference as well.

### Programming

We have to edit the data first. We wrote a function to wrangle our data.
After the call we want to know, weather the function has worked and if
there any improvement. That is why we make a quality control call again. We
can see that the function did exactly what we wanted it to do.

```{r Calculate_the_LRR, error=FALSE, message=FALSE, warning=FALSE}
# Load the function to prepare the data set
source("../../R/re_tidy.scripts/function.data.preparation.for.LRR.R")

# Function calls
DB_S1 <- data.preparation.for.LRR(raw.data)
# Function call
visualize.na.values.without.groups(DB_S1)
```

Now we calculate the LRR for each experiment...

```{r}
# Load the function to calculate the LRR by experiment
source("../../R/re_tidy.scripts/function.LRR.by.experiment.R")

# Function call to calculate the LRR for each experiment and save it as a HTML kable
LRR.by.experiment(DB_S1)
```

...and for each phase

```{r}
# Load the function to calculate the LRR by phase
source("../../R/re_tidy.scripts/function.LRR.by.phase.R")

# Function call to calculate the LRR for each phase and save it as a HTML kable
LRR.by.phase(DB_S1)
```

### Discussion

The result is rounded on 3 digits. As you can see, the results of Method 1
and Method 2 are different. In this case, Method 2 is more meaningful than
Method 1 (see theory). But this is not always the case. It is therefore
important to consider what exactly you want to analyze and what is the
context of the experiment.
